# Set up notebook
## Load libraries & functions
```{r}
source("libraries.R")
source("functions.R")
```

## Convert allele data in STRUCTURE format to ConStruct format

Following this guide: https://cran.r-project.org/web/packages/conStruct/vignettes/format-data.html#other-formats-to-construct

- copied populations.structure (structure file) to populations_structure.txt, as recommended
- then deleted first line from .txt file: "# Stacks v2.62;  Structure v2.3; June 10, 2023"

This is the "Allele frequency data".
```{r}
allele_frqs <- conStruct::structure2conStruct(
  infile = here::here("./clean_data/stacks_output/output_populations_no-IBD/mmaf0.05_R0.75/populations.structure.txt"),
  onerowperind = FALSE,
  start.loci = 3,
  start.samples = 2,
  missing.datum = -9,
  outfile = here("./clean_data/construct_data_20230613"))

head(allele_frqs)

```

## Load in other data
### Geographic sampling coordinates
#### (don't have to run)- Create coordinates csv
```{r}
# # coordinates of all populations
# all_coords <- read.csv(here::here("./clean_data/urb_metrics.csv"))
# 
# # the populations of this project, in order
# pop_map3_IDs <- read.csv(here::here("./genomic_resources/pop_map3.csv")) %>%
#   dplyr::rename("sample" = 1,
#                 "pop_id" = 2) %>%
# # add MW IDs to populations that are currently numbers
#   add_MW_IDs() 
# 
# # join coordinates w/current populations
# coords <- right_join(all_coords %>%
#                       dplyr::select(patch_id, lat, long),
#                     pop_map3_IDs %>%
#                       dplyr::select(-pop_id),
#                     by = "patch_id") %>%
#   arrange(sample) %>%
#   dplyr::select(long, lat) %>%
#   dplyr::rename("Lon" = 1,
#                 "Lat" = 2) %>%
#   as.matrix() %T>%
#   write.csv(here::here("./clean_data/conStruct_coords.csv"))
```

#### Load coordinates matrix
```{r}
coords <- read.csv(here::here("./clean_data/conStruct_coords.csv")) %>%
  dplyr::select(-X) %>%
  as.matrix()
```

### Geographic distance matrix
```{r}
# calculate pairwise great-circle distance between sampling coordinates
distances <- fields::rdist.earth(coords,
                            # distances in km
                            miles = FALSE)
```

# Run models (didn't actually run in RStudio- see note below)

At first, I ran these in Rstudio on my local computer. But then, I increased the required computational power and ran these on the HPC. These are example scripts.


## EXAMPLE spatial model: k = 1, 5,000 iterations, 3 chains
```{r}
#       the number of layers (K)
#       the allele frequency data (freqs)
#       the geographic distance matrix (geoDist)
#       the sampling coordinates (coords)

conStruct(spatial = TRUE, 
                    K = 1, 
                    freqs = allele_frqs,
                    geoDist = distances, 
                    coords = coords,
                    prefix = "sp_K1_iter5000_chains3",
                    n.chains = 3,
                    n.iter = 5000)
```

## EXAMPLE non-spatial model (like STRUCTURE): k = 1, 5,000 iterations, 3 chains
```{r}
#       the number of layers (K)
#       the allele frequency data (freqs)
#       the sampling coordinates (coords)
#
#   if you're running the nonspatial model, 
#       you do not have to specify 
#       the geographic distance matrix (geoDist)
#       
# options(mc.cores = 3)

conStruct(spatial = FALSE, 
                    K = 1, 
                    freqs = allele_frqs, 
                    geoDist = NULL, 
                    coords = coords,
                 #   prefix = "nsp_K1_iter5000_chains3",
                    prefix = "test",
                    n.chains = 3,
                    n.iter = 5,
          cores = 3)

parallel::stopCluster(3)

```

## Try parallelizing
### Non-spatial, k = 2
```{r}
library(foreach)
library(doParallel)

# Set the number of cores to be used
num_cores <- 6

# Initialize parallel backend using doParallel
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# Run non-spatial model in parallel
nonspatial_mod_run2 <- foreach(i = 1:num_cores, .packages = "conStruct") %dopar% {
  conStruct(spatial = FALSE,
            K = 2,
            freqs = allele_frqs,
            geoDist = NULL,
            coords = coords,
            prefix = paste0("nsp_K2_iter10000_chains5_", i),
            n.chains = 5,
            n.iter = 10000)
}

print("nonspatial mod run complete")

# Stop the parallel backend
stopCluster(cl)

```

# Cross-validation (didn't actually run in RStudio)
```{r}
##  k = 1:5, 10,000 iterations, 5 chains
x.validation(train.prop = 0.9,
                         n.reps = 10,
                         K = 1:5,
                         freqs = allele_frqs,
                         data.partitions = NULL,
                         geoDist = distances,
                         coords = coords,
                         prefix = "pilot",
                         n.iter = 10000,
                         make.figs = TRUE,
                         save.files = TRUE,
                         parallel = FALSE, #TRUE,
                         n.nodes = NULL) #5)
```

# Compare models
## Calculate layer contributions
### spatial
```{r}

# Create an empty matrix to store layer contributions
layer.contributions <- matrix(NA, nrow = 5, ncol = 5)

# Load the conStruct.results.Robj and data.block.Robj files for K=1
load( here::here("./conStruct/sp_k1/sp_K1_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/sp_k1/sp_K1_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=1
layer.contributions[, 1] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block), rep(0, 4))
tmp <- conStruct.results[[1]]$MAP$admix.proportions
tmp.order <- match.layers.x.runs(tmp,
                                 conStruct.results[[1]]$MAP$admix.proportions)  




# Load the conStruct.results.Robj and data.block.Robj files for K=2
load( here::here("./conStruct/sp_k2/sp_K2_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/sp_k2/sp_K2_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=2
layer.contributions[, 2] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block), rep(0,3))
tmp <- conStruct.results[[1]]$MAP$admix.proportions




# Load the conStruct.results.Robj and data.block.Robj files for K=3
load( here::here("./conStruct/sp_k3/sp_K3_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/sp_k3/sp_K3_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=3
layer.contributions[, 3] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block), rep(0, 2))
tmp <- conStruct.results[[1]]$MAP$admix.proportions
tmp.order <- match.layers.x.runs(tmp,
                                 conStruct.results[[1]]$MAP$admix.proportions)



# Load the conStruct.results.Robj and data.block.Robj files for K=4
load( here::here("./conStruct/sp_k4/sp_K4_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/sp_k4/sp_K4_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=4
layer.contributions[, 4] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block), rep(0, 1))
tmp <- conStruct.results[[1]]$MAP$admix.proportions
tmp.order <- match.layers.x.runs(tmp,
                                 conStruct.results[[1]]$MAP$admix.proportions)



# Load the conStruct.results.Robj and data.block.Robj files for K=5
load( here::here("./conStruct/sp_k5/sp_K5_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/sp_k5/sp_K5_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=5
layer.contributions[, 5] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block))
tmp <- conStruct.results[[1]]$MAP$admix.proportions
tmp.order <- match.layers.x.runs(tmp,
                                 conStruct.results[[1]]$MAP$admix.proportions)
```

### non-spatial
```{r}

# Create an empty matrix to store layer contributions
layer.contributions_nsp <- matrix(NA, nrow = 5, ncol = 5)

# Load the conStruct.results.Robj and data.block.Robj files for K=1
load( here::here("./conStruct/nsp_k1/nsp_K1_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/nsp_k1/nsp_K1_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=1
layer.contributions_nsp[, 1] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block), rep(0, 4))
tmp <- conStruct.results[[1]]$MAP$admix.proportions
tmp.order <- match.layers.x.runs(tmp,
                                 conStruct.results[[1]]$MAP$admix.proportions)  




# Load the conStruct.results.Robj and data.block.Robj files for K=2
load( here::here("./conStruct/nsp_k2/nsp_K2_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/nsp_k2/nsp_K2_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=2
layer.contributions_nsp[, 2] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block), rep(0,3))
tmp <- conStruct.results[[1]]$MAP$admix.proportions
tmp.order <- match.layers.x.runs(tmp,
                                 conStruct.results[[1]]$MAP$admix.proportions)


# Load the conStruct.results.Robj and data.block.Robj files for K=3
load( here::here("./conStruct/nsp_k3/nsp_K3_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/nsp_k3/nsp_K3_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=3
layer.contributions_nsp[, 3] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block), rep(0, 2))
tmp <- conStruct.results[[1]]$MAP$admix.proportions
tmp.order <- match.layers.x.runs(tmp,
                                 conStruct.results[[1]]$MAP$admix.proportions)



# Load the conStruct.results.Robj and data.block.Robj files for K=4
load( here::here("./conStruct/nsp_k4/nsp_K4_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/nsp_k4/nsp_K4_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=4
layer.contributions_nsp[, 4] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block), rep(0, 1))
tmp <- conStruct.results[[1]]$MAP$admix.proportions
tmp.order <- match.layers.x.runs(tmp, conStruct.results[[1]]$MAP$admix.proportions)

# not sure why, but I think two values are being swapped here (rows 2 and 4). I'm going to swap them manually so figure looks logical
new_row2 <- 0.968306016
new_row4 <- 0.001631156

layer.contributions_nsp[2,4] <- new_row2
layer.contributions_nsp[4,4] <- new_row4

# Load the conStruct.results.Robj and data.block.Robj files for K=5
load( here::here("./conStruct/nsp_k5/nsp_K5_iter10000_chains5_conStruct.results.Robj"))
load( here::here("./conStruct/nsp_k5/nsp_K5_iter10000_chains5_data.block.Robj"))

# Calculate layer contributions for K=5
layer.contributions_nsp[, 5] <- c(calculate.layer.contribution(conStruct.results[[1]], data.block))
tmp <- conStruct.results[[1]]$MAP$admix.proportions
tmp.order <- match.layers.x.runs(tmp,
                                 conStruct.results[[1]]$MAP$admix.proportions)

```

## Cross-validation
### read in non-spatial output
```{r}
nsp_k1 <- read.table(here::here("./results/conStruct/xval_k1/xval_nsp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 1)

nsp_k2 <- read.table(here::here("./results/conStruct/xval_k2_2/xval_nsp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 2)

nsp_k3 <- read.table(here::here("./results/conStruct/xval_k3/xval_nsp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 3)

nsp_k4 <- read.table(here::here("./results/conStruct/xval_k4/xval_nsp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 4)

nsp_k5 <- read.table(here::here("./results/conStruct/xval_k5/xval_nsp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 5)


nsp <- rbind(nsp_k1,
             nsp_k2,
             nsp_k3,
             nsp_k4,
             nsp_k5) %>%
  dplyr::mutate(type = "nsp") %>%
  rename_with(
    ~str_remove(., "nsp_"),
    starts_with("nsp_")
  )

nsp.CIs <- apply(nsp[, c(1:8)],1,function(x){mean(x) + c(-1.96,1.96) * sd(x)/length(x)})

```

### read in spatial output
```{r}
sp_k1 <- read.table(here::here("./results/conStruct/xval_k1/xval_sp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 1)

sp_k2 <- read.table(here::here("./results/conStruct/xval_k2_2/xval_sp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 2)


sp_k3 <- read.table(here::here("./results/conStruct/xval_k3/xval_sp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 3)


sp_k4 <- read.table(here::here("./results/conStruct/xval_k4/xval_sp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 4)

sp_k5 <- read.table(here::here("./results/conStruct/xval_k5/xval_sp_xval_results.txt"),
                     header=T, sep="\ ") %>%
  dplyr::mutate("k" = 5)

sp <- rbind(sp_k1,
            sp_k2,
            sp_k3,
            sp_k4,
            sp_k5) %>%
  dplyr::mutate(type = "sp") %>%
  rename_with(
    ~str_remove(., "sp_"),
    starts_with("sp_")
  )


# Combine both data frames
sp_nsp <- bind_rows(sp, nsp) %>%
  dplyr::select("type", "k", 1:8) %>%
  pivot_longer(cols = starts_with("rep_"),
               names_to = "replicate",
               values_to = "value")

# get CIs
sp_nsp_summary <- sp_nsp %>%
  group_by(type, k) %>%
  summarize(mean_value = mean(value),
            std_error = sd(value) / sqrt(n()))

# Plot the scatterplot with confidence intervals
ggplot(sp_nsp_summary,
       aes(x = k,
           y = mean_value,
           shape = type)) +
  geom_point(size = 3) +
  geom_errorbar(aes(
    ymin = mean_value - 1.96 * std_error, 
    ymax = mean_value + 1.96 * std_error),
                width = 0.1) +
  labs(x = "K",
       y = "Predictive accuracy",
       shape = "Model") +
  scale_shape_manual(values = c(16,1),
                     labels = c("Non-spatial", "Spatial")) +
  ggpubr::theme_pubr(legend = "right")

ggsave(plot = last_plot(),
       filename = "xvalidation.png",
       height = 8,
       width = 14,
       units = "cm",
       path = here::here("./Figures_Tables/conStruct/"))
```

# Export figures/tables
## Layer contribution plots
### Spatial
```{r}

png(filename = here::here("./Figures_Tables/conStruct/layer_contributions_spatial.png"),
     width = 1000, height = 800,
    units = "px",
    res = 225)

barplot(layer.contributions,
        col=c("blue", "red", "goldenrod1", "forestgreen", "darkorchid1"),
        xlab="",
        ylab="layer contributions",
        names.arg=paste0("K=",1:5))

dev.off()

# If a layer in a given model contributes very little to overall covariance, it is unlikely to have much biological significance. If you run conStruct analyses across values of K, and see that, after a certain value of K, no additional clusters contribute much to overall covariance, that may be a good indication that that value of K (or at least, no larger value of K) is best for describing the variation in your data.
```

### Non-spatial
```{r}

png(filename = here::here("./Figures_Tables/conStruct/layer_contributions_nonspatial.png"),
     width = 1000, height = 800,
    units = "px",
    res = 225)

barplot(layer.contributions_nsp,
        col=c("blue", "red", "goldenrod1", "forestgreen", "darkorchid1"),
        xlab="",
        ylab="layer contributions",
        names.arg=paste0("K=",1:5))

dev.off()
```

## Layer contribution tables
### Spatial
```{r}

layer.contributions %>%
  as.data.frame() %>%
  round(., digits = 3) %>%
  flextable::flextable() %>%
  flextable::set_header_labels(
    V1 = "K = 1", 
    V2 = "K = 2", 
    V3 = "K = 3", 
    V4 = "K = 4", 
    V5 = "K = 5" ) %>%
  flextable::save_as_docx(path = here::here("./Figures_Tables/conStruct/table_spatial.docx"))
```

### Non-spatial
```{r}

layer.contributions_nsp %>%
  as.data.frame() %>%
  round(., digits = 3) %>%
  flextable::flextable() %>%
  flextable::set_header_labels(
    V1 = "K = 1", 
    V2 = "K = 2", 
    V3 = "K = 3", 
    V4 = "K = 4", 
    V5 = "K = 5" ) %>%
  flextable::save_as_docx(path = here::here("./Figures_Tables/conStruct/table_nonspatial.docx"))

```


## Admixture plots
### One file
```{r}
# practice w/K=2, nonspatial model
load( here::here("./conStruct/nsp_k2/nsp_K2_iter10000_chains5_conStruct.results.Robj"))

# assign the MAP admixture proportions from the first MCMC chain to a variable with a new name
admix.props <- conStruct.results$chain_1$MAP$admix.proportions

# make a STRUCTURE plot using the maximum a posteriori (MAP) estimates from the first chain of a conStruct run

make.structure.plot(admix.proportions = admix.props)

# order plot by latitude
make.structure.plot(admix.proportions = admix.props,
                    sample.order = order(data.block$coords[,2]))


# to order the plot by dist to cc, I'll have to import those values from another csv and try to add them to the data.block list
urb <- read.csv(here("./clean_data/urb_metrics.csv")) %>%
  dplyr::rename("pop_id" = "patch_id") %>%
  dplyr::rename("Lon" = "long",
                "Lat" = "lat")

# join site coordinates with city_dist and pop ID values
urb_df <- data.block$coords %>%
  as.data.frame() %>%
  dplyr::left_join(.,
                   urb %>%
                     dplyr::select(City_dist,
                                   pop_id,
                                   Lat,
                                   Lon),
                   by = c("Lat", "Lon"))

# add city_dist values to data.block
data.block$dist <- urb_df %>%
  dplyr::select(City_dist) %>%
  as.matrix()


# order plot by city_dist
make.structure.plot(admix.proportions = admix.props,
                    sample.order = order(data.block$dist))



# add site names
# add pop_ID values to data.block
data.block$pop_id <- urb_df %>%
  dplyr::select(pop_id) %>%
  as.matrix()

# make it large enough to see all pop IDs
# png(filename = here::here("./Figures_Tables/conStruct/admixture_nonspatial_k2.png"),
#      width = 10000, height = 2000,
#     units = "px",
#     res = 250)

# see every few pop IDs and make y axis bigger
png(filename = here::here("./Figures_Tables/conStruct/admixture_nonspatial_k2_smaller.png"),
     width = 3000, height = 800,
    units = "px",
    res = 250)


# order plot by city_dist WITH pop names
make.structure.plot(admix.proportions = admix.props,
                    sample.order = order(data.block$dist),
                    sample.names = data.block$pop_id,
                    mar = c(4.5,4,2,2))


dev.off()

```

### All files
#### spatial
```{r}
# List of file paths for the "results.rObj" files
file_paths <- c(
  here::here("./conStruct/sp_k2/sp_K2_iter10000_chains5_conStruct.results.Robj"),
  here::here("./conStruct/sp_k3/sp_K3_iter10000_chains5_conStruct.results.Robj"),
  here::here("./conStruct/sp_k4/sp_K4_iter10000_chains5_conStruct.results.Robj") ,
  here::here("./conStruct/sp_k5/sp_K5_iter10000_chains5_conStruct.results.Robj")
)

# Loop over the file paths
for (file_path in file_paths) {
  # Load the results file
  load(file_path)
  
  # Assign the MAP admixture proportions from the first MCMC chain to a variable with a new name
  admix.props <- conStruct.results$chain_1$MAP$admix.proportions
  
  # Read urb_metrics.csv file
  urb <- read.csv(here("./clean_data/urb_metrics.csv")) %>%
    dplyr::rename("pop_id" = "patch_id") %>%
    dplyr::rename("Lon" = "long", "Lat" = "lat")
  
  # Join site coordinates with city_dist and pop ID values
  urb_df <- data.block$coords %>%
    as.data.frame() %>%
    dplyr::left_join(
      .,
      urb %>%
        dplyr::select(City_dist, pop_id, Lat, Lon),
      by = c("Lat", "Lon")
    )
  
  # Add city_dist values to data.block
  data.block$dist <- urb_df %>%
    dplyr::select(City_dist) %>%
    as.matrix()
  
  # Add pop_ID values to data.block
  data.block$pop_id <- urb_df %>%
    dplyr::select(pop_id) %>%
    as.matrix()

  # Extract the relevant part of the filename
  new_filename <- str_remove(basename(file_path), "_iter10000_chains5_conStruct.results")
  
  # Set the file name for saving the plot
  plot_filename <- here("./Figures_Tables/conStruct", paste0(new_filename, ".png"))

  # Create the plot
  png(filename = plot_filename, width = 3000, height = 800, units = "px", res = 250)
  
  make.structure.plot(
    admix.proportions = admix.props,
    sample.order = order(data.block$dist),
    sample.names = data.block$pop_id,
    mar = c(4.5, 4, 2, 2)
  )
  
  dev.off()
}

```


#### nonspatial
```{r}
# List of file paths for the "results.rObj" files
file_paths <- c(
  here::here("./conStruct/nsp_k2/nsp_K2_iter10000_chains5_conStruct.results.Robj"),
  here::here("./conStruct/nsp_k3/nsp_K3_iter10000_chains5_conStruct.results.Robj"),
  here::here("./conStruct/nsp_k4/nsp_K4_iter10000_chains5_conStruct.results.Robj"),
  here::here("./conStruct/nsp_k5/nsp_K5_iter10000_chains5_conStruct.results.Robj")
)

# Loop over the file paths
for (file_path in file_paths) {
  # Load the results file
  load(file_path)
  
  # Assign the MAP admixture proportions from the first MCMC chain to a variable with a new name
  admix.props <- conStruct.results$chain_1$MAP$admix.proportions
  
  # Read urb_metrics.csv file
  urb <- read.csv(here("./clean_data/urb_metrics.csv")) %>%
    dplyr::rename("pop_id" = "patch_id") %>%
    dplyr::rename("Lon" = "long", "Lat" = "lat")
  
  # Join site coordinates with city_dist and pop ID values
  urb_df <- data.block$coords %>%
    as.data.frame() %>%
    dplyr::left_join(
      .,
      urb %>%
        dplyr::select(City_dist, pop_id, Lat, Lon),
      by = c("Lat", "Lon")
    )
  
  # Add city_dist values to data.block
  data.block$dist <- urb_df %>%
    dplyr::select(City_dist) %>%
    as.matrix()
  
  # Add pop_ID values to data.block
  data.block$pop_id <- urb_df %>%
    dplyr::select(pop_id) %>%
    as.matrix()

  # Extract the relevant part of the filename
  new_filename <- str_remove(basename(file_path), "_iter10000_chains5_conStruct.results")
  
  # Set the file name for saving the plot
  plot_filename <- here("./Figures_Tables/conStruct", paste0(new_filename, ".png"))

  # Create the plot
  png(filename = plot_filename, width = 3000, height = 800, units = "px", res = 250)
  
  make.structure.plot(
    admix.proportions = admix.props,
    sample.order = order(data.block$dist),
    sample.names = data.block$pop_id,
    mar = c(4.5, 4, 2, 2)
  )
  
  dev.off()
}

```


## Extras
```{r}
# just for spatial K = 2
load("sp_K2_iter10000_chains5_conStruct.results.Robj")

admix.props <- conStruct.results$chain_1$MAP$admix.proportions

load("sp_K2_iter10000_chains5_data.block.Robj")
  
make.admix.pie.plot(admix.proportions = admix.props,
                        coords = data.block$coords,
                        radii = 0.8,
                        x.lim = c(-80.3,-78.8),
                        y.lim = c(43,44.4))
```


## Composite figure (xvalidation and layer contribution plot)
```{r}
xval_composite <- rasterGrob(readPNG(here::here("./Figures_Tables/conStruct/xvalidation.png")))

laycontr_composite <- rasterGrob(readPNG(here::here("./Figures_Tables/conStruct/layer_contributions_spatial.png")))


conStruct_composite_plots <- gridExtra::grid.arrange(xval_composite, laycontr_composite,
                        ncol = 2,
             left = textGrob("A", 
                            x = unit(1, "npc"), 
                            y = unit(0.95, "npc"),
                            gp = gpar(fontsize = 12)),
             right = textGrob("B", 
                            x = unit(-13, "npc"), 
                            y = unit(0.95, "npc"),
                            gp = gpar(fontsize = 12)))

ggsave("./Figures_Tables/conStruct/main_figs.png", 
       plot = conStruct_composite_plots,
       width = 6,
       height = 1.8*1.2,
       units = "in",
       dpi = 400)
```

