# Set up notebook
## load libraries
```{r}
source("libraries.R")
```

# Create sfs
## Load vcf2sfs code
```{r}
# can't install this package so copied the R code into my own file, load it

source(here::here("./scripts/vcf2sfs_code.R"))
```

## get path to vcf
```{r}
vcf_address <- here::here("./results/pi_fst_TajD_estimates/pi/global_pi/global_pi_vcf-all/populations.all.vcf")
```

## Notes for understanding vcfR::maf()'s output

*Each row is a locus.
*nAllele = the number of different alleles sequenced at this
 locus. At most, it's 2n = 261*2=522 because it's a diploid-
 so there's at most 2 alleles per locus for each individual.
 It can drop to numbers like 50 if only, for example,
 50/2n=50/2*261=50/522=~10% of the loci were sequenced.
*Count = the number of times a specific allele was seen at
 this locus.
*Frequency = Count/nAlleles to give the total frequency of
 that specific allele.

There can be bars in the 1 bins on the x axis because of
 missing values. Some loci were only sequenced in a few of
 the individuals (depending on R), so if an allele was seen
 only once but its locus was sequenced in only ~25% of
 individuals, it could still have a frequency >0.05 and
 wind up in the "1 individual" bin.

This holds up because the mmaf = 0.05, R = 1 plot should
 have no missing data and a min minor allele frequency of
 5%, so the first individual should show up at
 0.05*2n=0.05*2*261=26 individuals. It does.
 
## Read the VCF file and the popmap file and create a gt object
### SFS for all individuals in 1 population
```{r}
mygt <- vcf2gt(vcf_address,
               here::here("./genomic_resources/pop_map3_all1pop_for_sfs.txt"))

# look at all unique populations
unique(mygt$popmap)
```

### SFS for URBAN or RURAL individuals in 1 population
#### Urbanization = dist to CC
```{r}
mygt_urb <- vcf2gt(vcf_address,
               here::here("./clean_data/pop_maps/pop_map3_UR_dist.txt"))

# look at all unique populations
unique(mygt_urb$popmap)
```

#### Urbanization = urbanization score
```{r}
mygt_urb2 <- vcf2gt(vcf_address,
               here::here("./clean_data/pop_maps/pop_map3_UR_usc.txt"))

# look at all unique populations
unique(mygt_urb2$popmap)
```

## Create sfs
### SFS for all individuals in 1 population
```{r}
sfs <- gt2sfs.raw(mygt,
           "1") %>%
        as.data.frame() %>%
  dplyr::filter(X1 != 0) %T>%
  write_delim(here::here("./clean_data/sfs/sfs.txt"))

```

### SFS for URBAN individuals in 1 population
#### Urbanization = dist to CC
```{r}
sfs_urb <- gt2sfs.raw(mygt_urb,
           "Urban") %>%
        as.data.frame() %>%
  dplyr::filter(Urban != 0) %T>%
  write_delim(here::here("./clean_data/sfs/sfs_urban.txt"))

```

#### Urbanization = urbanization score
```{r}
sfs_urb2 <- gt2sfs.raw(mygt_urb2,
           "Urban") %>%
        as.data.frame() %>%
  dplyr::filter(Urban != 0) %T>%
  write_delim(here::here("./clean_data/sfs/sfs_urban_urbscore.txt"))

```

### SFS for RURAL individuals in 1 population
#### Urbanization = dist to CC
```{r}
sfs_rur <- gt2sfs.raw(mygt_urb,
           "Rural") %>%
        as.data.frame() %>%
  dplyr::filter(Rural != 0) %T>%
  write_delim(here::here("./clean_data/sfs/sfs_rural.txt"))

```

#### Urbanization = urbanization score
```{r}
sfs_rur2 <- gt2sfs.raw(mygt_urb2,
           "Rural") %>%
        as.data.frame() %>%
  dplyr::filter(Rural != 0) %T>%
  write_delim(here::here("./clean_data/sfs/sfs_rural_urbscore.txt"))

```

## Prepare to export sfs
### SFS for all individuals in 1 population
```{r}
# get freqs for SFS
freqs_for_stairwayplot <- sfs %>%
                   dplyr::select(Freq) %>%
  pull()


freqs_string <- paste(freqs_for_stairwayplot,
                      collapse = " ")

# Write the string to a txt file
writeLines(freqs_string,
           here::here("./clean_data/sfs/sfs_freqs.txt"))

# length of SFS
length(freqs_for_stairwayplot)
```

### SFS for URBAN individuals in 1 population
#### Urbanization = dist to CC
```{r}
# get freqs for SFS
sfs_urb %>%
  dplyr::select(Freq) %>%
  pull() %>%
  paste(collapse = " ") %T>%
# Write the string to a txt file
  writeLines(here::here("./clean_data/sfs/sfs_freqs_urb.txt"))

# length of SFS
length(sfs_urb %>%
  dplyr::select(Freq) %>%
  pull())
```

#### Urbanization = urbanization score
```{r}
# get freqs for SFS
sfs_urb2 %>%
  dplyr::select(Freq) %>%
  pull() %>%
  paste(collapse = " ") %T>%
# Write the string to a txt file
  writeLines(here::here("./clean_data/sfs/sfs_freqs_urb_urbscore.txt"))

# length of SFS
length(sfs_urb2 %>%
  dplyr::select(Freq) %>%
  pull())
```

### SFS for RURAL individuals in 1 population
#### Urbanization = dist to CC
```{r}
# get freqs for SFS
sfs_rur %>%
  dplyr::select(Freq) %>%
  pull() %>%
  paste(collapse = " ") %T>%
# Write the string to a txt file
  writeLines(here::here("./clean_data/sfs/sfs_freqs_rur.txt"))

# length of SFS
length(sfs_rur %>%
  dplyr::select(Freq) %>%
  pull())
```

#### Urbanization = urbanization score
```{r}
# get freqs for SFS
sfs_rur2 %>%
  dplyr::select(Freq) %>%
  pull() %>%
  paste(collapse = " ") %T>%
# Write the string to a txt file
  writeLines(here::here("./clean_data/sfs/sfs_freqs_rur_urbscore.txt"))

# length of SFS
length(sfs_rur2 %>%
  dplyr::select(Freq) %>%
  pull())
```
