# Set up notebook
## Load libraries & functions
```{r warning=F}
source(here::here("libraries.R"))
source(here::here("functions.R"))
```

## Import files
### Pop map 3
```{r}
# pop map 3 (compared to #1 has all individuals; #2 removed 2 low-quality individuals; #3 removed 5 individuals that were related to 5 others)
pop_map3 <- read.csv(
  here::here("./genomic_resources/pop_map3.csv"))  %>%
  dplyr::rename("pop_id" = "population") %>%
  tibble::rowid_to_column( "ID") %>%
  add_MW_IDs()

# urbanization data
urb <- read.csv(here("./clean_data/urb_metrics.csv")) 

# Join two datasets
full_popmap <- dplyr::left_join(pop_map3,
                         urb,
                         by = "patch_id" )
```

### Pop map 1
```{r}
# pop map 3 (compared to #1 has all individuals; #2 removed 2 low-quality individuals; #3 removed 5 individuals that were related to 5 others)
pop_map1 <- read.csv(
  here::here("./genomic_resources/original_sample_info/pop_map.csv"))  %>%
  dplyr::rename("pop_id" = "population") %>%
  tibble::rowid_to_column( "ID") %>%
  add_MW_IDs()

# Join two datasets
full_popmap1 <- dplyr::left_join(pop_map1,
                         urb,
                         by = "patch_id" )
```

# Export
## create popmap with sample, population, and urb/rur category
### Urb = dist to CC
#### Pop map 3
```{r}
full_popmap %>%
  dplyr::select(1:3, lat, long, 12:15) %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id, u_r_dist) %>%

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_all_dist.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)
```

#### Pop map 1 (all individuals... includes those that were later excluded due to low quality or high IBD)
```{r}
full_popmap1 %>%
  dplyr::select(1:3, lat, long, 12:15) %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id, u_r_dist) %>%

  write.csv(., 
            file = here::here("./clean_data/pop_maps/pop_map1_all_dist.csv"),
          #  stdout(),
          #  sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)
```

### Urb = urb score
#### Pop map 3
```{r}
full_popmap %>%
  dplyr::select(1:3, lat, long, 12:15) %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id, u_r_usc) %>%

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_all_usc.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)
```

#### Pop map 1 (all individuals... includes those that were later excluded due to low quality or high IBD)
```{r}
full_popmap1 %>%
  dplyr::select(1:3, lat, long, 12:15) %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id, u_r_usc) %>%

  write.csv(., 
            file = here::here("./clean_data/pop_maps/pop_map1_all_usc.csv"),
          #  stdout(),
          #  sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)
```


## for global fst and pi analyses
### urb = dist to CC
```{r}
full_popmap %>%
  dplyr::select(1:3, lat, long, 12:15) %>%
  arrange(ID) %>%
  dplyr::select(sample, u_r_dist) %>%

# urban vs. rural, determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_UR_dist.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

```


### urb = urb score
```{r}
full_popmap %>%
  dplyr::select(1:3, lat, long, 12:15) %>%
  arrange(ID) %>%
  dplyr::select(sample, u_r_usc) %>%

# urban vs. rural, determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_UR_usc.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

```

## for URBAN fst analyses
### Urb = dist to CC
#### pop map
```{r}
full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_dist) %>%
  dplyr::filter(u_r_dist == "Urban") %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id) %>%

# urban , determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_U_dist.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)



full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_dist) %>%
  dplyr::filter(u_r_dist == "Urban") %>%
  arrange(ID) %>%
  dplyr::select(sample) %>%

# urban , determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_UR_dist_URBAN.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

```

#### For vcftools fst: Each pop has its own txt 
```{r}

# Extract only urban pops
data <- full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_dist) %>%
  dplyr::filter(u_r_dist == "Urban") %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id) 

# Create a list to store the sample IDs for each population
populations <- unique(data$patch_id)
sample_lists <- vector("list", length(populations))

output_dir <- here("./clean_data/pop_maps/pop_maps_urban")

for (i in 1:length(populations)) {
  # Subset the data for the current population
  population <- populations[i]
  samples <- data[data$patch_id == population, "sample"]
  
  # Save the sample IDs to a file
  file_name <- paste0(population, ".txt")
  file_path <- file.path(output_dir, file_name)
  writeLines(as.character(samples), con = file_path)
}
```


#### For vcftools fst: Create part of vcftools fst command with populations
```{r}

file_names <- paste("pop_maps_urban", paste0(populations, ".txt"), sep = "/")

paste("--weir-fst-pop", paste(file_names, collapse = " --weir-fst-pop "), sep = " ")


```

### Urb = urb score
#### pop map
```{r}
full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_usc) %>%
  dplyr::filter(u_r_usc == "Urban") %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id) %>%

# urban , determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_U_usc.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)




full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_usc) %>%
  dplyr::filter(u_r_usc == "Urban") %>%
  arrange(ID) %>%
  dplyr::select(sample) %>%

# urban , determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_UR_usc_URBAN.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

```

#### For vcftools fst: Each pop has its own txt 
```{r}

# Extract only urban pops
data <- full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_usc) %>%
  dplyr::filter(u_r_usc == "Urban") %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id) 

# Create a list to store the sample IDs for each population
populations <- unique(data$patch_id)
sample_lists <- vector("list", length(populations))

output_dir <- here("./clean_data/pop_maps/pop_maps_urban_usc")

for (i in 1:length(populations)) {
  # Subset the data for the current population
  population <- populations[i]
  samples <- data[data$patch_id == population, "sample"]
  
  # Save the sample IDs to a file
  file_name <- paste0(population, ".txt")
  file_path <- file.path(output_dir, file_name)
  writeLines(as.character(samples), con = file_path)
}
```


#### For vcftools fst: Create part of vcftools fst command with populations
```{r}

file_names <- paste("pop_maps_urban_usc", paste0(populations, ".txt"), sep = "/")

paste("--weir-fst-pop", paste(file_names, collapse = " --weir-fst-pop "), sep = " ")


```


## for RURAL fst analyses
### Urb = dist to CC
#### pop map
```{r}
full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_dist) %>%
  dplyr::filter(u_r_dist == "Rural") %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id) %>%

# urban , determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_R_dist.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)




full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_dist) %>%
  dplyr::filter(u_r_dist == "Rural") %>%
  arrange(ID) %>%
  dplyr::select(sample) %>%

# urban , determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_UR_dist_RURAL.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

```

#### For vcftools fst: Each pop has its own txt 
```{r}

# Extract only urban pops
data <- full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_dist) %>%
  dplyr::filter(u_r_dist == "Rural") %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id) 

# Create a list to store the sample IDs for each population
populations <- unique(data$patch_id)
sample_lists <- vector("list", length(populations))

output_dir <- here("./clean_data/pop_maps/pop_maps_rural")

for (i in 1:length(populations)) {
  # Subset the data for the current population
  population <- populations[i]
  samples <- data[data$patch_id == population, "sample"]
  
  # Save the sample IDs to a file
  file_name <- paste0(population, ".txt")
  file_path <- file.path(output_dir, file_name)
  writeLines(as.character(samples), con = file_path)
}
```


#### For vcftools fst: Create part of vcftools fst command with populations
```{r}

file_names <- paste("pop_maps_rural", paste0(populations, ".txt"), sep = "/")

paste("--weir-fst-pop", paste(file_names, collapse = " --weir-fst-pop "), sep = " ")


```


### Urb = urb score
#### pop map
```{r}
full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_usc) %>%
  dplyr::filter(u_r_usc == "Rural") %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id) %>%

# urban , determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_R_usc.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)




full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_usc) %>%
  dplyr::filter(u_r_usc == "Rural") %>%
  arrange(ID) %>%
  dplyr::select(sample) %>%

# urban , determined by dist to CC

  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_UR_usc_RURAL.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

```

#### For vcftools fst: Each pop has its own txt 
```{r}

# Extract only urban pops
data <- full_popmap %>%
  dplyr::select(sample, ID, patch_id, u_r_usc) %>%
  dplyr::filter(u_r_usc == "Rural") %>%
  arrange(ID) %>%
  dplyr::select(sample, patch_id) 

# Create a list to store the sample IDs for each population
populations <- unique(data$patch_id)
sample_lists <- vector("list", length(populations))

output_dir <- here("./clean_data/pop_maps/pop_maps_rural_usc")

for (i in 1:length(populations)) {
  # Subset the data for the current population
  population <- populations[i]
  samples <- data[data$patch_id == population, "sample"]
  
  # Save the sample IDs to a file
  file_name <- paste0(population, ".txt")
  file_path <- file.path(output_dir, file_name)
  writeLines(as.character(samples), con = file_path)
}
```


#### For vcftools fst: Create part of vcftools fst command with populations
```{r}

file_names <- paste("pop_maps_rural_usc", paste0(populations, ".txt"), sep = "/")

paste("--weir-fst-pop", paste(file_names, collapse = " --weir-fst-pop "), sep = " ")

```

## for 1 individual/population (memgene, radish analyses)
### Urban & Rural both included
```{r}
# 122 unique populations
unique(pop_map3$pop_id) %>% length()

uniques <- pop_map3 %>%
  arrange(ID) %>%
  group_by(pop_id) %>%
  slice(1) %>%
  ungroup() %>% 
  arrange(ID) %>%
  dplyr::select(sample, pop_id)

uniques %>%
  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

uniques %>%
  write_csv(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop.csv"))
```

### Urban only
#### Urb = dist to CC
```{r}
uniques_urb <- full_popmap %>%
  arrange(ID) %>%
  group_by(pop_id) %>%
  slice(1) %>%
  ungroup() %>% 
  dplyr::filter(u_r_dist == "Urban") %>%
  arrange(ID) %>%
  dplyr::select(sample, pop_id)

uniques_urb %>%
  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop_URBAN.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

uniques_urb %>%
  write_csv(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop_URBAN.csv"))
```

#### Urb = urb score
```{r}
uniques_urb_usc <- full_popmap %>%
  arrange(ID) %>%
  group_by(pop_id) %>%
  slice(1) %>%
  ungroup() %>% 
  dplyr::filter(u_r_usc == "Urban") %>%
  arrange(ID) %>%
  dplyr::select(sample, pop_id)

uniques_urb_usc %>%
  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop_URBAN_usc.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

uniques_urb_usc %>%
  write_csv(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop_URBAN_usc.csv"))
```

### Rural only
#### Urb = dist to CC
```{r}
uniques_rur <- full_popmap %>%
  arrange(ID) %>%
  group_by(pop_id) %>%
  slice(1) %>%
  ungroup() %>% 
  dplyr::filter(u_r_dist == "Rural") %>%
  arrange(ID) %>%
  dplyr::select(sample, pop_id)

uniques_rur %>%
  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop_RURAL.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

uniques_rur %>%
  write_csv(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop_RURAL.csv"))
```


#### Urb = urb score
```{r}
uniques_rur_usc <- full_popmap %>%
  arrange(ID) %>%
  group_by(pop_id) %>%
  slice(1) %>%
  ungroup() %>% 
  dplyr::filter(u_r_usc == "Rural") %>%
  arrange(ID) %>%
  dplyr::select(sample, pop_id)

uniques_rur_usc %>%
  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop_RURAL_usc.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

uniques_rur_usc %>%
  write_csv(., 
            file = here::here("./clean_data/pop_maps/pop_map3_1indiv_per_pop_RURAL_usc.csv"))
```

## for all pops with >1 individual
### All individuals
```{r}
# find pops with 1 individual
pops_atleast1indiv <- pop_map3 %>%
  dplyr::group_by(pop_id) %>%
  dplyr::summarise(n_indiv = n()) %>%
  dplyr::filter(n_indiv > 1)

# join to keep only those pops
multi_indiv_pops <- dplyr::right_join(
  pop_map3,
  pops_atleast1indiv) %>%
  dplyr::arrange(ID) %>%
  dplyr::select(sample, pop_id)

multi_indiv_pops %>%
  write.table(., 
            file = here::here("./clean_data/pop_maps/pop_map3_MORE_THAN_1indiv_per_pop.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)

multi_indiv_pops %>%
  write_csv(., 
            file = here::here("./clean_data/pop_maps/pop_map3_MORE_THAN_1indiv_per_pop.csv"))
```

### Urban (~dist to CC)
```{r}
larger_urb_pops <- full_popmap %>%
  dplyr::right_join(., pops_atleast1indiv) %>%
  dplyr::select(sample, ID, pop_id, u_r_dist) %>%
  dplyr::filter(u_r_dist == "Urban") %>%
  arrange(ID) %>%
  dplyr::select(sample, pop_id)

# urban , determined by dist to CC

write.table(larger_urb_pops, 
            file = here::here("./clean_data/pop_maps/pop_map3_MORE_THAN_1indiv_per_pop_URBAN.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)


larger_urb_pops %>%
  write_csv(., 
            file = here::here("./clean_data/pop_maps/pop_map3_MORE_THAN_1indiv_per_pop_URBAN.csv"))

```

### Rural (~dist to CC)
```{r}
larger_rur_pops <- full_popmap %>%
  dplyr::right_join(., pops_atleast1indiv) %>%
  dplyr::select(sample, ID, pop_id, u_r_dist) %>%
  dplyr::filter(u_r_dist == "Rural") %>%
  arrange(ID) %>%
  dplyr::select(sample, pop_id)

# rural , determined by dist to CC

write.table(larger_rur_pops, 
            file = here::here("./clean_data/pop_maps/pop_map3_MORE_THAN_1indiv_per_pop_RURAL.txt"),
            stdout(),
            sep='\t',
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)


larger_rur_pops %>%
  write_csv(., 
            file = here::here("./clean_data/pop_maps/pop_map3_MORE_THAN_1indiv_per_pop_RURAL.csv"))

```