These analyses use the narrow, conservative genetic dataset:
-R = 0.75 & -min_maf = 0.05
-monomorphic loci excluded

# Set up notebook
## Load libraries & functions
```{r warning=F}
source(here::here("libraries.R"))
source(here::here("functions.R"))
```

# Import data
## Urbanization data
```{r}
urb <- read.csv(here::here("./clean_data/urb_metrics.csv")) %>%
  dplyr::mutate(site_id = as.factor(site_id),
                patch_id = as.factor(patch_id),
                urb_rur = as.factor(urb_rur),
                quadrant = as.factor(quadrant),
                river_valley = as.factor(river_valley)
                )
```

## Genetic data
```{r}
diversity_all_sites <- here::here("./clean_data/stacks_output/output_populations_no-IBD/mmaf0.05_R0.75/populations.sumstats_summary.tsv") %>%
  extract_variant_position_summary() %>%
# if population doesn't start with "MWI", it should start with "MW"
  # clean names
  janitor::clean_names() %>%
  # rename col 1
  dplyr::rename(pop_id = 1) %>%
  dplyr::mutate(pop_id = as.factor(pop_id)) %>%

### Add "MW" to numeric populations
  add_MW_IDs()

```

## Join genetic data with urbanization data
```{r}
# left vs. full join because some populations in urb_clean weren't genotyped due to lack of material
diversity_all_sites %<>%
  left_join(., urb, by = "patch_id") %>%
  dplyr::mutate(patch_id = as.factor(patch_id),
                pop_id = as.factor(pop_id))

```

# Does urbanization predict genetic diversity within pops? **R = 0.75, mmaf = 0.5**
## Explore data
### Structure
```{r}
str(diversity_all_sites)

head(diversity_all_sites)

diversity_all_sites %>%
  dplyr::select(1, 3, 4, "pi", "fis", "City_dist", "urb_score") %>%
  view

```

### Visualizations
```{r}
ggplot(diversity_all_sites, 
       aes(y= fis)) + 
  geom_boxplot()


# quick regressions of genetic vars against distance from cc
ggplot(diversity_all_sites, 
       aes(x = City_dist,
           y = fis)) + 
  geom_point()


qplot(data = diversity_all_sites,
      City_dist, num_indv)


# regress all selected variables against one another
diversity_all_sites %>%
  dplyr::select(pi, num_indv, fis, urb_score, City_dist) %>%
  pairs()


# look at distributions of urb vs rural pops
qplot(data = diversity_all_sites,
      u_r_dist, fis)

qplot(data = diversity_all_sites,
      u_r_usc, fis)

```

### Does urbanization correlate with number of individuals/population?
```{r}
test1 <- glmmTMB(City_dist ~ num_indv, diversity_all_sites)
performance::check_model(test1)
car::Anova(test1) # yes
r.squaredGLMM(test1) # only ~4% variance explained


test1 <- glmmTMB(urb_score ~ num_indv, diversity_all_sites)
performance::check_model(test1)
car::Anova(test1) # no
```

### Does the number of individuals per populations correlate with pop gen stats?
```{r}

ggplot(diversity_all_sites, 
       aes(x = num_indv,
           y = fis)) + 
  geom_point() +
  geom_smooth() +
  ggpubr::theme_pubr()

# this (and the df) shows us that populations with 1 individual all have fis = 0.
# I'll remove them from this analysis' dataset.



test1 <- glmmTMB(fis ~ num_indv,
                 diversity_all_sites %>%
                   dplyr::filter(fis != 0))
performance::check_model(test1)
car::Anova(test1) # yes
r.squaredGLMM(test1) # ~28% variance explained
# SO I WILL add num_indivs as a random effect to these models

```

## Urbanization is continuous
### Fit linear models/diagnostics
#### Inbreeding coefficient (FIS)
##### City_dist
```{r}

fis_dist <- glmmTMB(fis ~ City_dist + (1|num_indv),
           diversity_all_sites %>%
             dplyr::filter(fis != 0))

performance::check_model(fis_dist)
```

##### Urb_score
```{r}
fis_usc <- glmmTMB(fis ~ urb_score  + (1|num_indv),
           diversity_all_sites %>%
             dplyr::filter(fis != 0))

performance::check_model(fis_usc)
```

### ANOVA & effect sizes
```{r}

mod_list <- list(fis_dist,
                 fis_usc)


all_anovas <- do.call(rbind, create_anova_df(mod_list)) %>%
  view

# create flextable
all_anovas %>%
  dplyr::select(-Sig) %>%
  flextable::flextable() %>%
  flextable::compose(i = 1, j = 3, part = "header",
                     value = as_paragraph("Χ", as_sup("2"))) %>%
  flextable::compose(i = 1, j = 6, part = "header",
                     value = as_paragraph("R", as_sup("2"))) %>%
  bold(~ p <= 0.05) %>%
  flextable::autofit() %T>%
  flextable::save_as_docx(path = here::here("./Figures_Tables/regressions/fis_anova.docx"))
```

### Figures
#### FIS
##### City_dist
```{r}
fis_dist_pred <- ggeffects::ggpredict(fis_dist,
                                   terms = c("City_dist"),
                                   type = "fe")

fis_dist_pred_plot <- plot(fis_dist_pred, 
     add.data = F) +
  # geom_point(data = diversity_all_sites,
  #            aes(x = City_dist,
  #                y = fis)) +
  geom_ribbon(aes(x = x,
                  ymin = conf.low,
                  ymax = conf.high),
              # fill = "#66a182",
               alpha = 0.1
              ) +
  xlab("Distance to City Center (km)") +
  ylab("FIS") +
  ylim(0, 0.06) +
  labs(title = "") +
  theme_pubr() +
  labs_pubr() 

fis_dist_pred_plot
```

##### Urb_score
```{r}
fis_usc_pred <- ggeffects::ggpredict(fis_usc,
                                   terms = c("urb_score"),
                                   type = "fe")
 
fis_usc_pred_plot <- plot(fis_usc_pred, 
     add.data = F) +
  geom_ribbon(aes(x = x,
                  ymin = conf.low,
                  ymax = conf.high),
              # fill = "#66a182",
               alpha = 0.1
              ) +
  xlab("Urbanization Score") +
  ylab("FIS") +
  ylim(0, 0.06) +
  labs(title = "") +
  theme_pubr() +
  labs_pubr() +
  xlim(4, -4)

fis_usc_pred_plot
```

#### Number of individuals per pop
##### City_dist
```{r}
ggplot(diversity_all_sites,
       aes(x = City_dist,
           y = num_indv)) +
  xlab("Distance to City Center (km)") +
  ylab("Individuals/Population") +
  plot_aesthetics +
  theme_pubr() +
  labs_pubr() 

ggplot(diversity_all_sites,
       aes(x = num_indv)) +
  geom_histogram() +
  xlab("Individuals/Population") +
#  plot_aesthetics +
  theme_pubr() +
  labs_pubr() 
```

##### Urb_score
```{r}
ggplot(diversity_all_sites,
       aes(x = urb_score,
           y = num_indv)) +
  xlab("Urbanization Score") +
  ylab("Individuals/Population") +
  plot_aesthetics +
  theme_pubr() +
  labs_pubr() +
  xlim(4, -4)

```

#### Export all in one pdf
```{r}
# not including indivs/pop plots

(fis_dist_pred_plot | fis_usc_pred_plot) 

ggsave(here::here("./Figures_Tables/regressions/fis.png"),
       width = 10,
       height = 3.5,
       units = "in")
```

# Compute means per group, w/o accounting for diffs in pop size
## Urban
```{r}
# Urbanization = distance to city center
fis_u_dist <- diversity_all_sites %>%
             dplyr::filter(fis != 0) %>%
  dplyr::filter(u_r_dist == "Urban") %>%
  dplyr::select(fis) %>%
  colMeans() %>%
  round(4) %T>%
  print()

# Urbanization = urbanization score
fis_u_usc <- diversity_all_sites %>%
             dplyr::filter(fis != 0) %>%
  dplyr::filter(u_r_usc == "Urban") %>%
  dplyr::select(fis) %>%
  colMeans() %>%
  round(4) %T>%
  print()
```

## Rural
```{r}
# Urbanization = distance to city center
fis_r_dist <- diversity_all_sites %>%
             dplyr::filter(fis != 0) %>%
  dplyr::filter(u_r_dist == "Rural") %>%
  dplyr::select(fis) %>%
  colMeans() %>%
  round(4) %T>%
  print()

# Urbanization = urbanization score
fis_r_usc <- diversity_all_sites %>%
             dplyr::filter(fis != 0) %>%
  dplyr::filter(u_r_usc == "Rural") %>%
  dplyr::select(fis) %>%
  colMeans() %>%
  round(4) %T>%
  print()
```

## Put into table
```{r}
fis_summary <- data.frame(
  Urbanization = c("Distance to City Center", "Urbanization Score",
                   "Distance to City Center", "Urbanization Score"),
  Environment = c("Urban", "Urban",
                  "Rural", "Rural"),
  FIS = c(fis_u_dist, fis_u_usc,
          fis_r_dist, fis_r_usc))

fis_summary_wide <- fis_summary %>%
  pivot_wider(names_from = Environment, 
              values_from = FIS)

# Print the wide table
print(fis_summary_wide)
```

## Export table
```{r}
fis_summary_wide %>%
  flextable() %>%
  flextable::autofit() %>%
  save_as_docx(path = here::here("./Figures_Tables/fis/FIS_means.docx"))

```

