These analyses use the liberal genetic dataset:
-R = 0.5 & -min_maf = 2/256
-monomorphic loci included

# Set up notebook
## Load libraries & functions
```{r warning=F}
source(here::here("libraries.R"))
source(here::here("functions.R"))
```

# Import data
## Urbanization data
```{r}
urb <- read.csv(here::here("./clean_data/urb_metrics.csv")) %>%
  dplyr::mutate(site_id = as.factor(site_id),
                patch_id = as.factor(patch_id),
                urb_rur = as.factor(urb_rur),
                quadrant = as.factor(quadrant),
                river_valley = as.factor(river_valley)
                )
```

## Genetic data
```{r}
pi_data <- here::here("./clean_data/stacks_output/output_populations_no-IBD/mmaf0.007813_R0.5/output_populations_no-IBD-vcf-all/mmaf0.007813_R0.5/populations.sumstats_summary.tsv") %>%
  extract_all_position_summary() %>%
# if population doesn't start with "MWI", it should start with "MW"
  # clean names
  janitor::clean_names() %>%
  # rename col 1
  dplyr::rename(pop_id = 1) %>%
  dplyr::mutate(pop_id = as.factor(pop_id)) %>%

### Add "MW" to numeric populations
  add_MW_IDs()

```

## Join genetic data with urbanization data
```{r}
# left vs. full join because some populations in urb_clean weren't genotyped due to lack of material
pi_data %<>%
  left_join(., urb, by = "patch_id") %>%
  dplyr::mutate(patch_id = as.factor(patch_id),
                pop_id = as.factor(pop_id))

```

# Does urbanization predict genetic diversity within pops?
## Explore data
### Structure
```{r}
str(pi_data)

head(pi_data)

pi_data %>%
  dplyr::select(1, 3, 4, "pi", "fis", "City_dist", "urb_score") %>%
  view

```

### Visualizations
```{r}

ggplot(pi_data, 
       aes(y= pi)) + 
  geom_boxplot()


# quick regressions of genetic vars against distance from cc
ggplot(pi_data, 
       aes(x = City_dist,
           y = pi)) + 
  geom_point()

ggplot(pi_data, 
       aes(x = urb_score,
           y = pi)) + 
  geom_point() 

qplot(data = pi_data,
      City_dist, num_indv)


# regress all selected variables against one another
pi_data %>%
  dplyr::select(pi, num_indv, fis, urb_score, City_dist) %>%
  pairs()


# look at distributions of urb vs rural pops
qplot(data = pi_data,
      u_r_dist, pi)

qplot(data = pi_data,
      u_r_usc, pi)
```

### Does urbanization correlate with number of individuals/population?
```{r}
test1 <- glmmTMB(City_dist ~ num_indv, pi_data)
performance::check_model(test1)
car::Anova(test1) # yes
r.squaredGLMM(test1) # only ~4% variance explained


test1 <- glmmTMB(urb_score ~ num_indv, pi_data)
performance::check_model(test1)
car::Anova(test1) # no
```

### Does the number of individuals per populations correlate with pop gen stats?
#### Pi
```{r}
ggplot(pi_data %>%
                   dplyr::filter(num_indv > 1),
       aes(x = num_indv,
           y = pi)) +
  geom_smooth() +
  geom_point() +
  theme_pubr() +
  labs_pubr() 



test1 <- glmmTMB(pi ~ num_indv,
                 pi_data %>%
                   dplyr::filter(num_indv > 1))
performance::check_model(test1)
car::Anova(test1) # yes
r.squaredGLMM(test1) # ~34% variance explained
# SO I WILL add num_indivs as a fixed effect to these models AND remove pi for sample sites with 1 individual because in these cases, pi = observed heterozygosity

pi_data %>%
  dplyr::group_by(round(num_indv, 0)) %>%
  dplyr::summarise(pi_per_groupsize = mean(pi))

```

## Clean data
### Pi
```{r}
# remove sites with 1 individuals based on high variance explained, analysis in chunk above
pi_data %<>%
  dplyr::filter(num_indv > 1)
```

## Urbanization is continuous
### Fit linear models/diagnostics
#### Pi
##### City_dist
```{r}
pi_dist <- glmmTMB(pi ~ City_dist  + num_indv,
           pi_data)

performance::check_model(pi_dist)
car::Anova(pi_dist)


# % decrease with urb: (have to create pi_usc_pred object in Figures section below)
perc_decrease(pi_dist_pred) # 9.234%
```

##### Urb_score
###### All individuals
```{r}
pi_usc <- glmmTMB(pi ~ urb_score  + num_indv,
           pi_data)

performance::check_model(pi_usc)
car::Anova(pi_usc)

```

### ANOVA & effect sizes
```{r}

# urb DOES predict pi
dist_anova <- car::Anova(pi_dist)

# urb doesn't predict pi
usc_anova <- car::Anova(pi_usc) 

# export dist_anova
Rsq_dist <- MuMIn::r.squaredGLMM(pi_dist)[1] %>%
      round(digits = 3) %>%
      format(nsmall = 3) %T>%
  print
    
broom::tidy(dist_anova) %>%
      dplyr::mutate(
        "Variable" = "pi",
        "Rsq" = Rsq_dist
      ) %>%
      dplyr::rename(
        "Chi_sq" = statistic,
        "p" = p.value,
        "Urbanization" = term
      ) %>%
      dplyr::mutate(
        Chi_sq = format(round(Chi_sq, digits = 3), nsmall = 3),
        p = format(round(p, digits = 3), nsmall = 3)
      ) %>%
       dplyr::mutate(
         "p" = case_when(
           p <= 0.001 ~ "<0.001",
           TRUE ~ p
         )
      ) %>%
      dplyr::mutate(
        "Urbanization" = case_when(
          Urbanization == "City_dist" ~ "Distance",
          Urbanization == "urb_score" ~ "Urbanization Score",
          Urbanization == "num_indv" ~ "No. Individuals",
          TRUE ~ ""
        )) %>%
  flextable::flextable() %>%
  flextable::compose(i = 1, j = 2, part = "header",
                     value = as_paragraph("Χ", as_sup("2"))) %>%
  flextable::compose(i = 1, j = 6, part = "header",
                     value = as_paragraph("R", as_sup("2"))) %>%
  flextable::align(part = "all", j = c(2:6), align = "center") %>%
  # bold(~ p <= 0.05) %>%
  flextable::autofit() %T>%
  flextable::save_as_docx(path = here::here("./Figures_Tables/regressions/pi_anova_dist.docx"))


# export usc_anova

Rsq_usc <- MuMIn::r.squaredGLMM(pi_usc)[1] %>%
      round(digits = 3) %>%
      format(nsmall = 3) %T>%
  print
    
broom::tidy(usc_anova) %>%
      dplyr::mutate(
        "Variable" = "pi",
        "Rsq" = Rsq_usc
      ) %>%
      dplyr::rename(
        "Chi_sq" = statistic,
        "p" = p.value,
        "Urbanization" = term
      ) %>%
      dplyr::mutate(
        Chi_sq = format(round(Chi_sq, digits = 3), nsmall = 3),
        p = format(round(p, digits = 3), nsmall = 3)
      ) %>%
       dplyr::mutate(
         "p" = case_when(
           p <= 0.001 ~ "<0.001",
           TRUE ~ p
         )
      ) %>%
      dplyr::mutate(
        "Urbanization" = case_when(
          Urbanization == "City_dist" ~ "Distance",
          Urbanization == "urb_score" ~ "Urbanization Score",
          Urbanization == "num_indv" ~ "No. Individuals",
          TRUE ~ ""
        )) %>%
  flextable::flextable() %>%
  flextable::compose(i = 1, j = 2, part = "header",
                     value = as_paragraph("Χ", as_sup("2"))) %>%
  flextable::compose(i = 1, j = 6, part = "header",
                     value = as_paragraph("R", as_sup("2"))) %>%
  flextable::align(part = "all", j = c(2:6), align = "center") %>%
  # bold(~ p <= 0.05) %>%
  flextable::autofit() %T>%
  flextable::save_as_docx(path = here::here("./Figures_Tables/regressions/pi_anova_usc.docx"))
```

### Figures
#### Pi
##### City_dist
```{r}
pi_dist_pred <- ggeffects::ggpredict(pi_dist,
                                   terms = c("City_dist"),
                                   type = "fe")

pi_dist_pred_plot <- plot(pi_dist_pred, 
     add.data = F) +
  # add raw data
  geom_point(data = pi_data,
             aes(x = City_dist,
                 y = pi)) +
    geom_ribbon(aes(x = x,
                  ymin = conf.low,
                  ymax = conf.high),
              # fill = "#66a182",
               alpha = 0.1
              ) +
  xlab("Distance from the City Center (km)") +
  ylab("π") +
  scale_y_continuous(label = scales::comma,
                     limits = c(0.0006, 0.0009)) +
  labs(title = "") +
  theme_pubr() +
  labs_pubr() 

pi_dist_pred_plot
```

##### Urb_score
```{r}
pi_usc_pred <- ggeffects::ggpredict(pi_usc,
                                   terms = c("urb_score"),
                                   type = "fe")

pi_usc_pred_plot <- plot(pi_usc_pred, 
     add.data = F) +
    # add raw data
  geom_point(data = pi_data,
             aes(x = urb_score,
                 y = pi)) +
  geom_ribbon(aes(x = x,
                  ymin = conf.low,
                  ymax = conf.high),
              # fill = "#66a182",
               alpha = 0.1
              ) +
  xlab("Urbanization Score") +
  ylab("π") +
  scale_y_continuous(label = scales::comma,
                     limits = c(0.0006, 0.0009)) +  labs(title = "") +
  xlim(4, -4) +
  theme_pubr() +
  labs_pubr() 

pi_usc_pred_plot
```

#### Number of individuals per pop
##### City_dist
```{r}
ggplot(pi_data,
       aes(x = City_dist,
           y = num_indv)) +
  xlab("Distance to City Center (km)") +
  ylab("Individuals/Population") +
  plot_aesthetics +
  theme_pubr() +
  labs_pubr() 

ggplot(pi_data,
       aes(x = num_indv)) +
  geom_histogram() +
  xlab("Individuals/Population") +
#  plot_aesthetics +
  theme_pubr() +
  labs_pubr() 
```

##### Urb_score
```{r}
ggplot(pi_data,
       aes(x = urb_score,
           y = num_indv)) +
  xlab("Urbanization Score") +
  ylab("Individuals/Population") +
  plot_aesthetics +
  theme_pubr() +
  labs_pubr() +
  xlim(4, -4)

```

#### Export
```{r}

pi_dist_pred_plot + 
  pi_usc_pred_plot +
  patchwork::plot_layout(axes = "keep",
                         axis_titles = "collect_y") +
  plot_annotation(tag_levels = 'A')

ggsave(here::here("./Figures_Tables/regressions/pi.png"),
       width = 9,
       height = 3.5,
       units = "in")

```
