Using this tutorial as guidance: https://bookdown.org/hhwagner1/LandGenCourse_book/WE_4.html

# Set up notebook
## Load libraries & functions
```{r warning=F}
source(here::here("libraries.R"))
source(here::here("functions.R"))
```

## Import files
### All individuals
```{r}
# pop map with populations
pops <- read.csv(here("./genomic_resources/pop_map3.csv"))

n_unique_pops <- pops %>%
  dplyr::select(population) %>%
  dplyr::summarise(n_pops = unique(population)) %>%
  dplyr::summarise(n_pops = n()) %>%
  as.numeric() %T>% print

# pop map with urban/rural designations
pops_UR <- read.delim(here("./clean_data/pop_maps/pop_map3_UR_dist.txt"),
                      header = F) %>%
  dplyr::rename("sample" = 1,
                "urban_rural" = 2)

# join pop maps
pops <- full_join(pops,
                  pops_UR,
                  by = "sample")

# vcf
vcf <- read.vcfR(
  here::here("./clean_data/stacks_output/output_populations_no-IBD/mmaf0.05_R0.75/populations.snps.vcf"),
  verbose = FALSE)

## Convert vcf to genind object
my_genind <- vcfR2genind(vcf)


# add pops to genind
strata(my_genind) <- pops
pop(my_genind) <- pops[,"population"]
```

### Urban sites
```{r}
vcf_urb <- read.vcfR(
  here::here("./clean_data/filtered_vcfs/urban_pops_dist_mmaf0.05_variant_sites.recode.vcf"),
  verbose = FALSE)

## Convert vcf to genind object
my_genind_urb <- vcfR2genind(vcf_urb)


# read in pop map with populations
pops_urb <- read.csv(here("./clean_data/pop_maps/pop_map3_U_dist.csv"))

# add pops to genind
strata(my_genind_urb) <- pops_urb
pop(my_genind_urb) <- pops_urb[,"population"]
```

### Rural sites
```{r}
vcf_rur <- read.vcfR(
  here::here("./clean_data/filtered_vcfs/rural_pops_dist_mmaf0.05_variant_sites.recode.vcf"),
  verbose = FALSE)

## Convert vcf to genind object
my_genind_rur <- vcfR2genind(vcf_rur)


# read in pop map with populations
pops_rur <- read.csv(here("./clean_data/pop_maps/pop_map3_R_dist.csv"))

# add pops to genind
strata(my_genind_rur) <- pops_rur
pop(my_genind_rur) <- pops_rur[,"population"]
```

### Sites with >1 individual
```{r}
vcf_larger_pops <- read.vcfR(
  here::here("./clean_data/filtered_vcfs/pops_MORE_THAN_1indiv_per_pop_mmaf0.05_variant_sites.recode.vcf"),
  verbose = FALSE)

## Convert vcf to genind object
my_genind_larger_pops <- vcfR2genind(vcf_larger_pops)



# read in pop map with ALL populations
pops_all <- read.csv(here("./genomic_resources/pop_map3.csv"))

# read in pop map with ONLY populations >1 indiv
pops_larger <- read.csv(here("./clean_data/pop_maps/pop_map3_MORE_THAN_1indiv_per_pop.csv")) %>%
  dplyr::rename(population = pop_id) %>%
  dplyr::mutate(population = as.character(population))

# pop map with urban/rural designations
pops_UR <- read.delim(here("./clean_data/pop_maps/pop_map3_UR_dist.txt"),
                      header = F) %>%
  dplyr::rename("sample" = 1,
                "urban_rural" = 2)



# join pop maps
pops_larger_pops <- full_join(pops_all,
                  pops_UR,
                  by = "sample") %>%
  # keep only pops with >1 indiv
  dplyr::right_join(.,
                    pops_larger,
                    by = c("sample", "population"))



# add pops to genind
strata(my_genind_larger_pops) <- pops_larger_pops
pop(my_genind_larger_pops) <- pops_larger_pops[,"population"]
```

# AMOVA- Analysis of Molecular Variance
## All individuals
```{r}
# estimate the percent of molecular variance at each hierarchical level
amova.result <- poppr::poppr.amova(my_genind,
                hier = ~urban_rural/population, 
                within=FALSE, # specifies that within-individual variance (i.e., observed heterozygosity) should not be tested. Setting this to ‘TRUE’ can lead to problems with missing values.
                method = "ade4")

amova.result

# test whether each variance component is statistically significant (i.e., significantly larger than zero)
amova.test <- ade4::randtest(amova.result, nrepet = 999)

amova.test

# plot
plot(amova.test)
```

## Urban individuals
```{r}
# estimate the percent of molecular variance at each hierarchical level
amova.urban <- poppr::poppr.amova(my_genind_urb,
                hier = ~population, 
                within=FALSE, 
                method = "ade4")

amova.urban


# $call
# ade4::amova(samples = xtab, distances = xdist, structures = xstruct)
# 
# $results
#                  Df    Sum Sq  Mean Sq
# Between samples  61  4408.614 72.27236
# Within samples   88  6522.134 74.11516
# Total           149 10930.749 73.36073
# 
# $componentsofcovariance
#                                  Sigma          %
# Variations  Between samples -0.7673451  -1.046173
# Variations  Within samples  74.1151647 101.046173
# Total variations            73.3478196 100.000000
# 
# $statphi
#                           Phi
# Phi-samples-total -0.01046173




# test whether variance component is statistically significant (i.e., significantly larger than zero)
amova.test_urb <- ade4::randtest(amova.urban, nrepet = 999)

amova.test_urb

```

## Rural individuals
```{r}
# estimate the percent of molecular variance at each hierarchical level
amova.rural <- poppr::poppr.amova(my_genind_rur,
                hier = ~population, 
                within=FALSE,
                method = "ade4")

amova.rural

# $call
# ade4::amova(samples = xtab, distances = xdist, structures = xstruct)
# 
# $results
#                  Df   Sum Sq  Mean Sq
# Between samples  59 4045.743 68.57191
# Within samples   46 3168.011 68.86981
# Total           105 7213.754 68.70242
# 
# $componentsofcovariance
#                                  Sigma           %
# Variations  Between samples -0.1705773  -0.2482958
# Variations  Within samples  68.8698096 100.2482958
# Total variations            68.6992323 100.0000000
# 
# $statphi
#                            Phi
# Phi-samples-total -0.002482958






# test whether variance component is statistically significant (i.e., significantly larger than zero)
amova.test_rur <- ade4::randtest(amova.rural, nrepet = 999)

amova.test_rur

```

## Sites with >1 individual
```{r}
amova.larger_pops <- poppr::poppr.amova(my_genind_larger_pops,
                hier = ~urban_rural/population, 
                within=FALSE,
                method = "ade4")

amova.larger_pops

# $call
# ade4::amova(samples = xtab, distances = xdist, structures = xstruct)
# 
# $results
#                                     Df      Sum Sq  Mean Sq
# Between urban_rural                  1    69.37201 69.37201
# Between samples Within urban_rural  43  3164.14861 73.58485
# Within samples                     134  9462.89198 70.61860
# Total                              178 12696.41260 71.32816
# 
# $componentsofcovariance
#                                                      Sigma            %
# Variations  Between urban_rural                -0.05657655  -0.07933814
# Variations  Between samples Within urban_rural  0.74864234   1.04983226
# Variations  Within samples                     70.61859686  99.02950588
# Total variations                               71.31066264 100.00000000
# 
# $statphi
#                                   Phi
# Phi-samples-total        0.0097049412
# Phi-samples-urban_rural  0.0104900000
# Phi-urban_rural-total   -0.0007933814




# test whether variance component is statistically significant (i.e., significantly larger than zero)
amova.test_larger_pops <- ade4::randtest(amova.larger_pops, nrepet = 999)

amova.test_larger_pops

```

# Export
## General AMOVA
```{r}
amova_general <- amova.result$results %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename("Group" = 1,
                "df" = 2,
                "SS" = 3,
                "MS" = 4) %>%
  dplyr::mutate(across("SS": "MS",
                       ~ round(., 3))) %>%
  dplyr::mutate(Group = case_when(
         Group == "Between urban_rural  " ~ "Among urban/rural groups",
         Group == "Between samples Within urban_rural" ~ "Among populations within urban/rural groups",
         Group == "Within samples" ~ "Within populations",
         TRUE ~ Group
     ))

amova_general %>%
  flextable() %>%
  flextable::align(j = c(2:4), align = "center", part = "all") %>%
  flextable::autofit() %T>%
  flextable::save_as_docx(path = here::here(
    "./Figures_Tables/amova/amova.docx"))
```

## Components of covariance
```{r}
amova_cov <- amova.result$componentsofcovariance %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename("Group" = 1,
                "Variation" = 3) %>%
  dplyr::mutate(Group = str_replace(Group,
                                        "Variations ", ""),
                Group = str_replace(Group,
                                        " variations", "")) %>%
  dplyr::mutate(across(Sigma:Variation,
                       ~ round(., 3))) %>%
  dplyr::mutate(Variation = paste0(Variation, "%")) %>%
  dplyr::mutate(Group = case_when(
         Group == " Between urban_rural  " ~ "Among urban/rural groups",
         Group == " Between samples Within urban_rural" ~ "Among populations within urban/rural groups",
         Group == " Within samples" ~ "Within populations",
         TRUE ~ Group
     )) 

amova_cov %>%
  flextable() %>%
  flextable::align(j = c(2:3), align = "center", part = "all") %>%
  flextable::autofit() %T>%
  flextable::save_as_docx(path = here::here(
    "./Figures_Tables/amova/amova_covariance.docx"))
```

## Phi
```{r}
amova_phi <- amova.result$statphi %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  dplyr::rename("Group" = 1) %>%
  dplyr::mutate(across(Phi, ~ round(., 3))) %>%
  dplyr::mutate(Group = case_when(
         Group == "Phi-samples-total" ~ "Populations:Total",
         Group == "Phi-samples-urban_rural" ~ "Populations:Urban/rural groups",
         Group == "Phi-urban_rural-total" ~ "Urban/rural groups:Total",
         TRUE ~ Group
     ))

# table with full names of groups
amova_phi %>%
  flextable() %>%
  flextable::align(j = 2, align = "center", part = "all") %>%
  flextable::autofit() %T>%
  flextable::save_as_docx(path = here::here(
    "./Figures_Tables/amova/amova_phi.docx"))


# table with abbreviations for groups
## Individuals = I; Populations = P; Groups = G
amova_phi %>%
  dplyr::mutate(Group = case_when(
         Group == "Populations:Total" ~ "PT",
         Group == "Populations:Urban/rural groups" ~ "PG",
         Group == "Urban/rural groups:Total" ~ "GT",
         TRUE ~ Group
     )) %>%
  flextable() %>%
  flextable::align(j = 2, align = "center", part = "all") %>%
  flextable::autofit() %T>%
  flextable::save_as_docx(path = here::here(
    "./Figures_Tables/amova/amova_phi_abbrevs.docx"))
```

## P-values
```{r}
# have to recreate amova table to export it as a table
amova_list <- list(
  amova.test$obs,
  amova.test$alter,
  amova.test$names,
  amova.test$expvar$Std.Obs,
  amova.test$pvalue
  )

list_a <- purrr::map(amova_list,
                     tibble::as_tibble)

names(list_a) <- c("obs", "alter", "names", "Std.obs", "p")

list2env(list_a, envir = .GlobalEnv)

amova_p <- cbind(names, obs, Std.obs, alter, p) %>%
  dplyr::rename(Test = 1,
                Obs = 2,
                Std.Obs = 3,
                Alternative = 4,
                p = 5) %>%
  dplyr::mutate(Test = case_when(
         Test == "Variations within samples" ~ "Variation within populations",
         Test == "Variations between samples" ~ "Variation between populations",
         Test == "Variations between urban_rural" ~ "Variation between urban/rural groups",
         TRUE ~ Test
     )) %>%
    dplyr::mutate(across(c(Obs:Std.Obs, p),
                       ~ round(., 3)))

amova_p %>%
  flextable() %>%
  flextable::align(j = c(2:5), align = "center", part = "all") %>%
  flextable::italic(j = 5, part = "header") %>%
  flextable::autofit() %T>%
  flextable::save_as_docx(path = here::here(
    "./Figures_Tables/amova/amova_significance.docx"))
  

```


## Combine the above tables (except phi) into one table
```{r}
amova_all <- amova_cov %>%
  dplyr::full_join(amova_general, .) %>%
  dplyr::full_join(.,
                   amova_p %>%
                     dplyr::select(Test, p) %>%
                     dplyr::rename("Group" = 1) %>%
                     dplyr::mutate(Group = case_when(
         Group == "Variation within populations" ~ "Within populations",
         Group == "Variation between populations" ~ "Among populations within urban/rural groups",
         Group == "Variation between urban/rural groups" ~ "Among urban/rural groups",
         TRUE ~ Group
     )))

amova_all %>%
  flextable() %>%
  flextable::align(j = 2:7, align = "center", part = "all") %>%
  flextable::italic(j = 7, part = "header") %>%
  flextable::autofit() %T>%
  flextable::save_as_docx(path = here::here(
    "./Figures_Tables/amova/amova_combined_table.docx"))
  
```

## Plot
```{r}
png(here::here("./Figures_Tables/amova/test.png"),
    width = 12,
    height = 12,
    units = "cm",
    res = 200)

plot(amova.test)

dev.off()


```

