# Set up notebook
## Load packages
```{r}
library(tidyverse)
library(here)
library(magrittr)
library(janitor)
library(geosphere)
```

# Import data
```{r}
# urbanization data
urb <- read.csv(here::here("./raw_data/102_Sites_updatedASlatlongs.csv")) %>%
  janitor::clean_names() %>%
  dplyr::select(-c(6,7,8))

# urbanization score data
usc <- read.csv(here::here("./raw_data/Urb_index_sgs.csv")) %>%
  janitor::clean_names()
```

# Join dist to CC with urb score data
```{r}
urb_clean <- full_join(urb,
                       usc,
                       by = c("site_id",
                              "patch_id",
                              "lat",
                              "long"))

```

# Get haversine distances from populations to city center
```{r}
# Ref lat and longs are for Yonge & Dundas intersection in downtown Toronto
urb_clean$Ref_Lat <- 43.656327
urb_clean$Ref_Long <- -79.380904


# Find distances from Yonge/Dundas to sample sites (in meters)
urb_clean %<>%
  dplyr::mutate(CTD_m = geosphere::distHaversine(cbind(long, lat),
                                      cbind(Ref_Long, Ref_Lat))) %>%
  # convert to km
  dplyr::mutate(City_dist = CTD_m / 1000) %>%
  # drop ref lat, long, and city_dist (in m) cols
  dplyr::select(-c(Ref_Lat, Ref_Long, CTD_m))
```

# Reclassify the NAs from that prior site as urb/rur based on distance from CC and urb_score.
make new columns: u_r_dist & u_r_usc (urban or rural)
this column is different from urb_rur, which I made when scouting sites for the 100 populations. I don't have those types of designations for the sites I imported from my previous observational study (note the NAs). So I'll calculate two new columns that assign a population an "urban" or "rural" status based on its distance from city center (u_r_dist) or urbanization score (u_r_usc)
```{r}
urb_clean %<>%
  dplyr::mutate(u_r_dist = case_when(
    City_dist <= 30 ~ "Urban",
    TRUE ~ "Rural")) %>%
  dplyr::mutate(u_r_usc = case_when(
    urb_score > 0 ~ "Urban",
    TRUE ~ "Rural"))
```

# Export
```{r}
write.csv(urb_clean,
          here::here("./clean_data/urb_metrics.csv"))
```