---
title: How does urbanization predict pi and FIS?
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---
TO DO:
- make sure y axis is transformed for figures
- export figures

These analyses use the narrow, conservative genetic dataset:
-R = 1 & -min_maf = 0.05

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Set up notebook
## Load libraries & functions
```{r warning=F}
source(here::here("libraries.R"))
source(here::here("functions.R"))
```

## Import data
### Urbanization data
```{r}
urb <- read.csv(here::here("./clean_data/urb_metrics.csv")) %>%
  dplyr::mutate(site_id = as.factor(site_id),
                patch_id = as.factor(patch_id),
                urb_rur = as.factor(urb_rur),
                quadrant = as.factor(quadrant),
                river_valley = as.factor(river_valley)
                )
```

### Genetic data
```{r}

diversity_variant_sites <- here::here("./summary_stats/mmaf0.05_R1_populations_output/populations.sumstats_summary.tsv") %>%
  extract_variant_position_summary() %>%
# if population doesn't start with "MWI", it should start with "MW"
  # clean names
  janitor::clean_names() %>%
  # rename col 1
  dplyr::rename(pop_id = 1) %>%
  dplyr::mutate(pop_id = as.factor(pop_id)) %>%

### Add "MW" to numeric populations
  add_MW_IDs()

```

## Join genetic data with urbanization data
```{r}
# left vs. full join because some populations in urb_clean weren't genotyped due to lack of material
diversity_variant_sites %<>%
  left_join(., urb, by = "patch_id") %>%
  dplyr::mutate(patch_id = as.factor(patch_id),
                pop_id = as.factor(pop_id))


# make new columns: u_r_dist & u_r_usc (urban or rural)
# this column is different from urb_rur, which I made when scouting sites for the 100 populations. I don't have those types of designations for the sites I imported from my previous observational study (note the NAs). So I'll calculate two new columns that assign a population an "urban" or "rural" status based on its distance from city center (u_r_dist) or urbanization score (u_r_usc)
diversity_variant_sites %<>%
  dplyr::mutate(u_r_dist = case_when(
    City_dist <= 40 ~ "Urban",
    TRUE ~ "Rural")) %>%
  dplyr::mutate(u_r_usc = case_when(
    urb_score > 0 ~ "Urban",
    TRUE ~ "Rural"))

```

# Does urbanization predict genetic diversity within pops?
## Explore data
### Structure
```{r}
str(diversity_variant_sites)

head(diversity_variant_sites)

diversity_variant_sites %>%
  dplyr::select(1, 3, 4, "pi", "fis", "City_dist", "urb_score") %>%
  view

```

### Visualizations
```{r}
ggplot(diversity_variant_sites, 
       aes(y= fis)) + 
  geom_boxplot()

ggplot(diversity_variant_sites, 
       aes(y= pi)) + 
  geom_boxplot()


# quick regressions of genetic vars against distance from cc
ggplot(diversity_variant_sites, 
       aes(x = City_dist,
           y = fis)) + 
  geom_point()

ggplot(diversity_variant_sites, 
       aes(x = City_dist,
           y = pi)) + 
  geom_point()

qplot(data = diversity_variant_sites,
      City_dist, num_indv)


# regress all selected variables against one another
diversity_variant_sites %>%
  dplyr::select(pi, num_indv, fis, urb_score, City_dist) %>%
  pairs()


# look at distributions of urb vs rural pops
qplot(data = diversity_variant_sites,
      u_r_dist, pi)

qplot(data = diversity_variant_sites,
      u_r_usc, pi)

qplot(data = diversity_variant_sites,
      u_r_dist, fis)

qplot(data = diversity_variant_sites,
      u_r_usc, fis)

```

### Does urbanization correlate with number of individuals/population?
```{r}
test1 <- glmmTMB(City_dist ~ num_indv, diversity_variant_sites)
performance::check_model(test1)
car::Anova(test1) # yes
r.squaredGLMM(test1) # only ~3% variance explained
```

### Does the number of individuals per populations correlate with pop gen stats?
#### FIS
```{r}

ggplot(diversity_variant_sites, 
       aes(x = num_indv,
           y = fis)) + 
  geom_point(aes(color = City_dist)) +
  scale_color_continuous(type = "viridis")

# this (and the df) shows us that populations with 1 individual all have fis = 0.
# I'll remove them from this analysis' dataset.



test1 <- glmmTMB(fis ~ num_indv,
                 diversity_variant_sites %>%
                   dplyr::filter(fis != 0))
performance::check_model(test1)
car::Anova(test1) # no
r.squaredGLMM(test1) # ~1% variance explained
# SO I won't add num_indivs as a random effect to these models

```

#### Pi
```{r}
qplot(diversity_variant_sites$num_indv,
      diversity_variant_sites$pi)


test1 <- glmmTMB(pi ~ num_indv,
                 diversity_variant_sites)
performance::check_model(test1)
car::Anova(test1) # no
# SO I WON'T add num_indivs as a random effect to these models
```

## Urbanization is continuous
### Fit linear models/diagnostics
#### Inbreeding coefficient (FIS)
##### City_dist
```{r}

fis_dist <- glmmTMB(fis ~ City_dist,
           diversity_variant_sites %>%
             dplyr::filter(fis != 0))

performance::check_model(fis_dist)
```

##### Urb_score
```{r}
fis_usc <- glmmTMB(fis ~ urb_score ,
           diversity_variant_sites %>%
             dplyr::filter(fis != 0))

performance::check_model(fis_usc)
```

#### Pi
##### City_dist
```{r}
pi_dist <- glmmTMB(pi ~ City_dist ,
           diversity_variant_sites)

performance::check_model(pi_dist)
```

##### Urb_score
```{r}
pi_usc <- glmmTMB(pi ~ urb_score ,
           diversity_variant_sites)

performance::check_model(pi_usc)
```

### ANOVA & effect sizes
```{r}

mod_list <- list(fis_dist,
                 fis_usc,
                 pi_dist,
                 pi_usc)


all_anovas <- do.call(rbind, create_anova_df) %>%
  view

# create flextable
all_anovas %>%
  dplyr::select(-Sig) %>%
  flextable::flextable() %>%
  flextable::compose(i = 1, j = 3, part = "header",
                     value = as_paragraph("Χ", as_sup("2"))) %>%
  flextable::compose(i = 1, j = 6, part = "header",
                     value = as_paragraph("R", as_sup("2"))) %>%
  bold(~ p <= 0.05) %>%
  flextable::autofit()
```

### Figures
#### FIS
##### City_dist
```{r}
fis_dist_pred <- ggeffects::ggpredict(fis_dist,
                                   terms = c("City_dist"),
                                   type = "fe")

fis_dist_pred_plot <- plot(fis_dist_pred, 
     add.data = F) +
  geom_ribbon(aes(x = x,
                  ymin = conf.low,
                  ymax = conf.high),
              # fill = "#66a182",
               alpha = 0.1
              ) +
  xlab("Distance to City Center (km)") +
  ylab("FIS") +
  ylim(-0.06, 0) +
  labs(title = "") +
  theme_pubr() +
  labs_pubr() 

fis_dist_pred_plot
```

##### Urb_score
```{r}
fis_usc_pred <- ggeffects::ggpredict(fis_usc,
                                   terms = c("urb_score"),
                                   type = "fe")
 
fis_usc_pred_plot <- plot(fis_usc_pred, 
     add.data = F) +
  geom_ribbon(aes(x = x,
                  ymin = conf.low,
                  ymax = conf.high),
              # fill = "#66a182",
               alpha = 0.1
              ) +
  xlab("Urbanization Score") +
  ylab("FIS") +
  ylim(-0.06, 0) +
  labs(title = "") +
  theme_pubr() +
  labs_pubr() +
  xlim(4, -4)

fis_usc_pred_plot
```

#### Pi
##### City_dist
```{r}
pi_dist_pred <- ggeffects::ggpredict(pi_dist,
                                   terms = c("City_dist"),
                                   type = "fe")

pi_dist_pred_plot <- plot(pi_dist_pred, 
     add.data = F) +
  geom_ribbon(aes(x = x,
                  ymin = conf.low,
                  ymax = conf.high),
              # fill = "#66a182",
               alpha = 0.1
              ) +
  xlab("Distance to City Center (km)") +
  ylab("Pi") +
  ylim(0.22, 0.32) +
  labs(title = "") +
  theme_pubr() +
  labs_pubr() 

pi_dist_pred_plot
```

###### Urb_score
```{r}
pi_usc_pred <- ggeffects::ggpredict(pi_usc,
                                   terms = c("urb_score"),
                                   type = "fe")

pi_usc_pred_plot <- plot(pi_usc_pred, 
     add.data = F) +
  geom_ribbon(aes(x = x,
                  ymin = conf.low,
                  ymax = conf.high),
              # fill = "#66a182",
               alpha = 0.1
              ) +
  xlab("Urbanization Score") +
  ylab("Pi") +
  ylim(0.22, 0.32) +
  labs(title = "") +
  xlim(4, -4) +
  theme_pubr() +
  labs_pubr() 

pi_usc_pred_plot
```

#### Number of individuals per pop
##### City_dist
```{r}
ggplot(diversity_variant_sites,
       aes(x = City_dist,
           y = num_indv)) +
  xlab("Distance to City Center (km)") +
  ylab("Individuals/Population") +
  plot_aesthetics +
  theme_pubr() +
  labs_pubr() 

```

##### Urb_score
```{r}
ggplot(diversity_variant_sites,
       aes(x = urb_score,
           y = num_indv)) +
  xlab("Urbanization Score") +
  ylab("Individuals/Population") +
  plot_aesthetics +
  theme_pubr() +
  labs_pubr() +
  xlim(4, -4)

```

#### Export all in one pdf
```{r}
# not including indivs/pop plots

(fis_dist_pred_plot | fis_usc_pred_plot) /
(pi_dist_pred_plot | pi_usc_pred_plot)

ggsave(here::here("./Figures_Tables/regressions/pi_fis.png"),
       width = 10,
       height = 7,
       units = "in")

```

## Urbanization is categorical
### Fit linear models/diagnostics
#### Inbreeding coefficient (FIS)
##### Distance
```{r}
fis_dist_cat <- glmmTMB(fis ~ u_r_dist,
           diversity_variant_sites %>%
             dplyr::filter(fis != 0))

performance::check_model(fis_dist_cat)
```

##### Urb score
```{r}
fis_usc_cat <- glmmTMB(fis ~ u_r_usc,
           diversity_variant_sites %>%
             dplyr::filter(fis != 0))

performance::check_model(fis_usc_cat)
```

#### Pi
##### Distance
```{r}
pi_dist_cat <- glmmTMB(pi ~ u_r_dist ,
           diversity_variant_sites)

performance::check_model(pi_dist_cat)
```

##### Urb score
```{r}
pi_usc_cat <- glmmTMB(pi ~ u_r_usc,
           diversity_variant_sites)

performance::check_model(pi_usc_cat)
```

### ANOVA & effect sizes
```{r}

mod_list_cat <- list(
                 fis_dist_cat,
                 fis_usc_cat,
                 pi_dist_cat,
                 pi_usc_cat)

all_anovas_cat <- do.call(rbind, create_anova_df_cat) %>%
  view

# create flextable
all_anovas_cat %>%
  dplyr::select( -Sig) %>%
  flextable::flextable() %>%
  flextable::compose(i = 1, j = 3, part = "header",
                     value = as_paragraph("Χ", as_sup("2"))) %>%
  flextable::compose(i = 1, j = 6, part = "header",
                     value = as_paragraph("R", as_sup("2"))) %>%
  bold(~ p <= 0.05) %>%
  flextable::autofit()
```

### Figures
#### FIS
##### City_dist
```{r}

fis_dist_cat_pred <- ggeffects::ggpredict(fis_dist_cat,
                                   terms = c("u_r_dist"),
                                   type = "fe")

fis_dist_cat_pred_plot <- ggplot(data = fis_dist_cat_pred,
         aes(x = x, y = predicted)) +
        geom_point(color = "black",
                   size = 2) + 
    geom_errorbar(aes(ymin = conf.low,
                      ymax = conf.high),
                  width = 0.2,
                  color = "black") +
  ylim(-0.06, 0) +
  xlab("Urbanization (based on\nDistance to City Center)") +
  ylab("FIS") +
  labs(title = "") +
  theme_pubr() +
  labs_pubr()

fis_dist_cat_pred_plot
```

##### Urb_score
```{r}

fis_usc_cat_pred <- ggeffects::ggpredict(fis_usc_cat,
                                   terms = c("u_r_usc"),
                                   type = "fe")

fis_usc_cat_pred_plot <- ggplot(data = fis_usc_cat_pred,
         aes(x = x, y = predicted)) +
        geom_point(color = "black",
                   size = 2) + 
    geom_errorbar(aes(ymin = conf.low,
                      ymax = conf.high),
                  width = 0.2,
                  color = "black") +
  ylim(-0.06, 0) +
  xlab("Urbanization (based on\nUrbanization Score)") +
  ylab("FIS") +
  labs(title = "") +
  theme_pubr() +
  labs_pubr() 

fis_usc_cat_pred_plot
```

#### Pi
##### City_dist
```{r}

pi_dist_cat_pred <- ggeffects::ggpredict(pi_dist_cat,
                                   terms = c("u_r_dist"),
                                   type = "fe")

pi_dist_cat_pred_plot <- ggplot(data = pi_dist_cat_pred,
         aes(x = x, y = predicted)) +
        geom_point(color = "black",
                   size = 2) + 
    geom_errorbar(aes(ymin = conf.low,
                      ymax = conf.high),
                  width = 0.2,
                  color = "black") +
  ylim(0.25, 0.3) +
  xlab("Urbanization (based on\nDistance to City Center)") +
  ylab("Pi") +
  labs(title = "") +
  theme_pubr() +
  labs_pubr() 

pi_dist_cat_pred_plot
```

###### Urb_score
```{r}
pi_usc_cat_pred <- ggeffects::ggpredict(pi_usc_cat,
                                   terms = c("u_r_usc"),
                                   type = "fe")

pi_usc_cat_pred_plot <- ggplot(data = pi_usc_cat_pred,
         aes(x = x, y = predicted)) +
        geom_point(color = "black",
                   size = 2) + 
    geom_errorbar(aes(ymin = conf.low,
                      ymax = conf.high),
                  width = 0.2,
                  color = "black") +
  ylim(0.25, 0.3) +
  xlab("Urbanization (based on\nUrbanization Score)") +
  ylab("Pi") +
  labs(title = "") +
  theme_pubr()  +
  labs_pubr() 

pi_usc_cat_pred_plot
```

#### Export all in one pdf
```{r}
(fis_dist_cat_pred_plot | fis_usc_cat_pred_plot) /
(pi_dist_cat_pred_plot | pi_usc_cat_pred_plot )

ggsave(here::here("./Figures_Tables/regressions/fis_pi_categorical.png"),
       width = 10,
       height = 7,
       units = "in")

```