# Set up notebook
## load libraries
```{r}
source("libraries.R")
library(cowplot)
```

## Load vcf2sfs code
```{r}
# can't install this package so copied the R code into my own file, load it

source(here::here("./R_scripts/exploration/vcf2sfs_code.R"))
```

## get path to vcf
```{r}
vcf_address <- here::here("./pi_fst_TajD_estimates/pi/global_pi_vcf-all/populations.all.vcf")

```

## Notes for understanding vcfR::maf()'s output

*Each row is a locus.
*nAllele = the number of different alleles sequenced at this
 locus. At most, it's 2n = 261*2=522 because it's a diploid-
 so there's at most 2 alleles per locus for each individual.
 It can drop to numbers like 50 if only, for example,
 50/2n=50/2*261=50/522=~10% of the loci were sequenced.
*Count = the number of times a specific allele was seen at
 this locus.
*Frequency = Count/nAlleles to give the total frequency of
 that specific allele.

There can be bars in the 1 bins on the x axis because of
 missing values. Some loci were only sequenced in a few of
 the individuals (depending on R), so if an allele was seen
 only once but its locus was sequenced in only ~25% of
 individuals, it could still have a frequency >0.05 and
 wind up in the "1 individual" bin.

This holds up because the mmaf = 0.05, R = 1 plot should
 have no missing data and a min minor allele frequency of
 5%, so the first individual should show up at
 0.05*2n=0.05*2*261=26 individuals. It does.
 
# Read the VCF file and the popmap file and create a gt object
## SFS for all individuals in 1 population
```{r}
mygt <- vcf2gt(vcf_address,
               here::here("./genomic_resources/pop_map3_all1pop_for_sfs.txt"))

# look at all unique populations
unique(mygt$popmap)
```

## SFS for URBAN or RURAL individuals in 1 population
### Urbanization = dist to CC
```{r}
mygt_urb <- vcf2gt(vcf_address,
               here::here("./clean_data/pop_map3_UR_dist.txt"))

# look at all unique populations
unique(mygt_urb$popmap)
```

### Urbanization = urbanization score
```{r}
mygt_urb2 <- vcf2gt(vcf_address,
               here::here("./clean_data/pop_map3_UR_usc.txt"))

# look at all unique populations
unique(mygt_urb2$popmap)
```

# Create sfs
## SFS for all individuals in 1 population
```{r}
sfs <- gt2sfs.raw(mygt,
           "1") %>%
        as.data.frame() %>%
  dplyr::filter(X1 != 0) %T>%
  write_delim(here::here("./StairwayPlot/sfs.txt"))

```

## SFS for URBAN individuals in 1 population
### Urbanization = dist to CC
```{r}
sfs_urb <- gt2sfs.raw(mygt_urb,
           "Urban") %>%
        as.data.frame() %>%
  dplyr::filter(Urban != 0) %T>%
  write_delim(here::here("./StairwayPlot/sfs_urban.txt"))

```

### Urbanization = urbanization score
```{r}
sfs_urb2 <- gt2sfs.raw(mygt_urb2,
           "Urban") %>%
        as.data.frame() %>%
  dplyr::filter(Urban != 0) %T>%
  write_delim(here::here("./StairwayPlot/sfs_urban_urbscore.txt"))

```

## SFS for RURAL individuals in 1 population
### Urbanization = dist to CC
```{r}
sfs_rur <- gt2sfs.raw(mygt_urb,
           "Rural") %>%
        as.data.frame() %>%
  dplyr::filter(Rural != 0) %T>%
  write_delim(here::here("./StairwayPlot/sfs_rural.txt"))

```

### Urbanization = urbanization score
```{r}
sfs_rur2 <- gt2sfs.raw(mygt_urb2,
           "Rural") %>%
        as.data.frame() %>%
  dplyr::filter(Rural != 0) %T>%
  write_delim(here::here("./StairwayPlot/sfs_rural_urbscore.txt"))

```

# Prepare to export sfs
## SFS for all individuals in 1 population
```{r}
# get freqs for SFS
freqs_for_stairwayplot <- sfs %>%
                   dplyr::select(Freq) %>%
  pull()


freqs_string <- paste(freqs_for_stairwayplot,
                      collapse = " ")

# Write the string to a txt file
writeLines(freqs_string,
           here::here("./StairwayPlot/sfs_freqs.txt"))

# length of SFS
length(freqs_for_stairwayplot)
```

## SFS for URBAN individuals in 1 population
### Urbanization = dist to CC
```{r}
# get freqs for SFS
sfs_urb %>%
  dplyr::select(Freq) %>%
  pull() %>%
  paste(collapse = " ") %T>%
# Write the string to a txt file
  writeLines(here::here("./StairwayPlot/sfs_freqs_urb.txt"))

# length of SFS
length(sfs_urb %>%
  dplyr::select(Freq) %>%
  pull())
```

### Urbanization = urbanization score
```{r}
# get freqs for SFS
sfs_urb2 %>%
  dplyr::select(Freq) %>%
  pull() %>%
  paste(collapse = " ") %T>%
# Write the string to a txt file
  writeLines(here::here("./StairwayPlot/sfs_freqs_urb_urbscore.txt"))

# length of SFS
length(sfs_urb2 %>%
  dplyr::select(Freq) %>%
  pull())
```

## SFS for RURAL individuals in 1 population
### Urbanization = dist to CC
```{r}
# get freqs for SFS
sfs_rur %>%
  dplyr::select(Freq) %>%
  pull() %>%
  paste(collapse = " ") %T>%
# Write the string to a txt file
  writeLines(here::here("./StairwayPlot/sfs_freqs_rur.txt"))

# length of SFS
length(sfs_rur %>%
  dplyr::select(Freq) %>%
  pull())
```

### Urbanization = urbanization score
```{r}
# get freqs for SFS
sfs_rur2 %>%
  dplyr::select(Freq) %>%
  pull() %>%
  paste(collapse = " ") %T>%
# Write the string to a txt file
  writeLines(here::here("./StairwayPlot/sfs_freqs_rur_urbscore.txt"))

# length of SFS
length(sfs_rur2 %>%
  dplyr::select(Freq) %>%
  pull())
```

# Stairway Plot figures
## Create plot-exporting function
```{r}
make_ne_plot <- function(df, exported_filepath){
  
  # full plot without red horizontal line at 500 years ago
  full_plot <- ggplot(
    df,
    aes(
      x = year,
      y = Ne_median))  +
    geom_segment( # add red vertical line at 500 yrs
      aes(
        x = 500,
        xend = 500,
        y = 0,
        yend = max(df$Ne_median*1.1)),
      color = "red",
      linetype = 2 ) +
    geom_point(size = 0.75) +
    scale_x_continuous(labels = scales::comma) +
    scale_y_log10(breaks=c(1, 10, 100, 1000, 10000, 100000),
                  labels = scales::comma,
                  limits = c(1, NA)) +
    xlab("Years before present") +
    ylab("Median Ne") +
    theme_bw() +
    labs_pubr() +
    theme(plot.margin = unit(c(0.5,1,0.5,0.5), "cm"))
  
  ### Plot inset (past 500 years)
  full_inset <- full_plot +
    scale_x_continuous(limits = c(NA, 500),
                       labels = scales::comma) +
    theme(axis.title = element_blank(),
          plot.background = element_rect(
            colour = "black", fill = "white", size = 1))
  
  ### Overlay inset in full figure
  cowplot::ggdraw() +
    draw_plot(full_plot) +
    draw_plot(full_inset,
              x = 0.45,
              y = 0.15, 
              width = 0.45,
              height = 0.5)
  
  
  #### Export
  ggsave(filename = here::here(exported_filepath),
         height = 5,
         width = 7,
         dpi = 200)
}

```

## All sites in 1 population
```{r}
full_fig <- read.delim(here::here("./StairwayPlot/run_full_20230908/milkweed_20230908.final.summary")) %>%
  dplyr::select(-c(theta_per_site_median,
                   theta_per_site_2.5.,
                   theta_per_site_97.5.,
                   n_estimation,
                   mutation_per_site))

head(full_fig)

make_ne_plot(full_fig,
             "./Figures_Tables/Stairway_plot/full_with_inset.png")
```

## Urban
### Urb = distance to city center
```{r}
urb_fig <- read.delim(here::here("./StairwayPlot/run_urban_20230911/milkweed_urban_dist.final.summary")) %>%
  dplyr::select(-c(theta_per_site_median,
                   theta_per_site_2.5.,
                   theta_per_site_97.5.,
                   n_estimation,
                   mutation_per_site))

head(urb_fig)


make_ne_plot(urb_fig,
             "./Figures_Tables/Stairway_plot/urban_with_inset.png")
```

### Urb = urbanization score
```{r}
urb_fig2 <- read.delim(here::here("./StairwayPlot/run_urban_usc/milkweed_urban_urbscore.final.summary")) %>%
  dplyr::select(-c(theta_per_site_median,
                   theta_per_site_2.5.,
                   theta_per_site_97.5.,
                   n_estimation,
                   mutation_per_site))

head(urb_fig2)


make_ne_plot(urb_fig2,
             "./Figures_Tables/Stairway_plot/urban_with_inset_urbscore.png")
```

## Rural
### Urb = distance to city center
```{r}
rur_fig <- read.delim(here::here("./StairwayPlot/run_rural_20230911/milkweed_rural_dist.final.summary")) %>%
  dplyr::select(-c(theta_per_site_median,
                   theta_per_site_2.5.,
                   theta_per_site_97.5.,
                   n_estimation,
                   mutation_per_site))

head(rur_fig)


make_ne_plot(rur_fig,
             "./Figures_Tables/Stairway_plot/rural_with_inset.png")
```

### Urb = urbanization score
```{r}
rur_fig2 <- read.delim(here::here("./StairwayPlot/run_rural_usc/milkweed_rural_urbscore.final.summary")) %>%
  dplyr::select(-c(theta_per_site_median,
                   theta_per_site_2.5.,
                   theta_per_site_97.5.,
                   n_estimation,
                   mutation_per_site))

head(rur_fig2)


make_ne_plot(rur_fig2,
             "./Figures_Tables/Stairway_plot/rural_with_inset_urbscore.png")

```
